---
title: "OA-ICC Data Download and Curation"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

This notebook has code used to fetch and QA/QC data from from the OA-ICC Database for Holm et al. For code that compares and assigns biome conditions to treatments see 'OA-ICC Biome Compare'

Load Packages and set working directory. 
```{r}
# needed packages - install if necessary 

library(dplyr) # data reshaping
library(stringr) # string tools
library(readr) # read files
library(purrr) #  functional programming tools
library(pangaear) # access PANGAEA data
library(tidyr) # data reshaping
library(reticulate) # using python functions
library(ggplot2) # plotting 

library(here) # set wd at cloned repo

```

Get data chunk.
Fetch Data from PANGAEA (rate limited rate limited on PANGAEA side)
```{r}
# import dois 
loc <- here("data/OA_ICC_Data/OA_ICC_Study_selection.xlsx")
selection <- readxl::read_xlsx(path = loc, sheet = "Worksheet_header")
dois <- paste0('10.1594/',selection$Id)

# create python env
virtualenv_create("pangaeapy-env")
use_virtualenv("pangaeapy-env", required = TRUE)
virtualenv_install("pangaeapy-env",packages = c("pangaeapy","pandas"))

# import pangaeapy and pandas
reticulate::py_require("pangaeapy")
pangaeapy <- import("pangaeapy")
pd <- import("pandas")

# function to fetch a dataset by DOI
import_dataset <- function(doi) {
  print(doi)
  # create PanDataSet object
  ds <- pangaeapy$PanDataSet(doi)
  
  # convert Python data frame to R
  dat <- py_to_r(ds$data)
  
  # keep parameters as Python objects for full info
  params <- ds$parameters
  
  list(
    doi        = doi,
    data       = dat,
    parameters = params
  )
}

datasets <- lapply(dois,import_dataset) # import

remove(dois,import_dataset,loc,pangaeapy,pd,selection) # clean up

```

QA / QC Chunk 
This section QA and QC's OA-ICC data. The final output is 'OA_ICC' which matches data in OA_ICC_data_merged.csv
```{r}

# clean parameter info
slots <- names(datasets[[1]]$parameters$`Study type`) # get slots

empty_like <- function(x) {
  if (is.list(x)) {
    map(x, empty_like)  # recurse on sub-lists
  } else {
    list()              # replace non-list values with empty list
  }
}

# get a clean framework
parameters_clean <- map(datasets,~ .x[["parameters"]]) %>% empty_like()

# clean parameter data and make it an R object 
for (data in 1:length(datasets)) {
  for (param in names(datasets[[data]][["parameters"]])) {
    for (slot in slots) {
      x <- datasets[[data]][["parameters"]][[param]][[slot]]
      if ("python.builtin.method" %in% class(x)) {
        parameters_clean[[data]][[param]][[slot]] <- NA
      }else{
        parameters_clean[[data]][[param]][[slot]] <- x
      }
    }
  }
}

# replace
for (data in 1:length(datasets)) {
  datasets[[data]][["parameters"]] <-  parameters_clean[[data]]
}

# further reformat sub objects
for (data in 1:length(datasets)) {
  for (param in names(datasets[[data]][["parameters"]])) {
    for (slot in names(datasets[[data]][["parameters"]][[param]])) {
      x <- datasets[[data]][["parameters"]][[param]][[slot]]
      if ("python.builtin.object" %in% class(x)) {
        slts <- names(datasets[[data]][["parameters"]][[param]][[slot]])
        replace <- list()
        for (slt in slts) {
          replace[[slt]] <- datasets[[data]][["parameters"]][[param]][[slot]][[slt]]
        }
        datasets[[data]][["parameters"]][[param]][[slot]] <- replace
      }else{
        datasets[[data]][["parameters"]][[param]][[slot]] <- x
      }
    }
  }
}

# explicitly drop studies that did not return a data.frame (drops 5 studies)
df_lines <- as.numeric(datasets %>% map(~.x$data %>% nrow))
sum(df_lines == 0) # 5
datasets <- datasets[which(df_lines != 0)]
length(datasets)

# get columns where the methods match "Calculated using seacarb after Nisumaa et al. (2010)"
# OR it was inputed by Yan, Yang. He didn't all add this methods tag.
# Meaning it was calculated after the fact for OA-ICC

for (data in 1:length(datasets)) {
  for (param in names(datasets[[data]][["parameters"]])) {
    # set false
    datasets[[data]][["carb_params"]][[param]] <- FALSE

    # check method
    x <- datasets[[data]][["parameters"]][[param]][["method"]][["name"]]
    if (!is.null(x)) {
      if (x == "Calculated using seacarb after Nisumaa et al. (2010)") {
        datasets[[data]][["carb_params"]][[param]] <- TRUE
      }
    }
    
    # check PI
    x <- datasets[[data]][["parameters"]][[param]][["PI"]][['name']]
    if (!is.null(x)) {
      if (x == "Yang, Yan") {
        datasets[[data]][["carb_params"]][[param]] <- TRUE
      }
    }
  }
}

carb_check <- datasets %>% map(~ {
  tot <- unlist(.x$carb_params)
  data.frame(hits = sum(tot), total = length(tot), study = .x$doi)
}) %>% bind_rows() 

sum(carb_check$hits == 0)

# for treatment params
for (data in 1:length(datasets)) {
  for (param in names(datasets[[data]][["parameters"]])) {
    # set false
    datasets[[data]][["treat_params"]][[param]] <- FALSE
    
    # check long name
    x <- datasets[[data]][["parameters"]][[param]][["name"]]
    if (!is.null(x)) {
      if (str_detect(x,regex("treat", ignore_case = TRUE))) {
        datasets[[data]][["treat_params"]][[param]] <- TRUE
      }
    }
    # check short name
    x <- datasets[[data]][["parameters"]][[param]][["shortName"]]
    if (!is.null(x)) {
      if (str_detect(x,regex("treat", ignore_case = TRUE))) {
        datasets[[data]][["treat_params"]][[param]] <- TRUE
      }
    }
  }
}

treat_check <- datasets %>% map(~ {
  tot <- unlist(.x$treat_params)
  data.frame(hits = sum(tot), total = length(tot), study = .x$doi)
}) %>% bind_rows() 

sum(treat_check$hits == 0)

# if we didn't have any treatment column, make CO2 ion conc the treatment 
no_treat_no_co2 <- c()
for (data in 1:length(datasets)) {
  if((treat_check$hits == 0)[data]){
    w <- which(unlist(datasets[[data]]$carb_params) & str_detect(names(datasets[[data]]$carb_params),"^CO2$|^CO2_2$|^CO2_3$"))
    if(length(w) == 0){
        print(c("ERROR in",data))
        no_treat_no_co2 <- c(no_treat_no_co2,data)
      }else{
        datasets[[data]]$treat_params[w[1]] <- TRUE # set at true in treat params
    }
  }
}

# for temperature params
for (data in 1:length(datasets)) {
  for (param in names(datasets[[data]][["parameters"]])) {
    # set false
    datasets[[data]][["temp_params"]][[param]] <- FALSE
    
    # check short name
    x <- datasets[[data]][["parameters"]][[param]][["shortName"]]
    if (!is.null(x)) {
      if (str_detect(x,regex("temp", ignore_case = TRUE))) {
        datasets[[data]][["temp_params"]][[param]] <- TRUE
      }
    }
  }
}

temp_check <- datasets %>% map(~ {
  tot <- unlist(.x$temp_params)
  data.frame(hits = sum(tot), total = length(tot), study = .x$doi)
}) %>% bind_rows() 

sum(temp_check$hits == 0)

# also pull any column that includes "treat" or "treatment" in short or long name
# NOTE - get long name treatment columns too
carb_only <- map(datasets, function(ds) {
  doi <- ds$doi
  
  # columns listed in carb_params
  sel_cols <- names(ds$carb_params)[unlist(ds$carb_params)]
  
  # pull comments safely as character vector
  sel_cols_comment <- vapply(
    ds$parameters[unlist(ds$carb_params)],
    function(x) {
      comment <- if (!is.null(x$comment)) x$comment else ""
      # remove "total scale" and commas
      comment <- gsub("total scale", "", comment, ignore.case = TRUE)
      comment <- gsub(", ?", "", comment)
      # trim whitespace
      comment <- trimws(comment)
      # if empty, set to "No Comment"
      if (comment == "") "No Comment" else comment
    },
    FUN.VALUE = character(1)
  )
  names(sel_cols_comment) <- sel_cols
  
  # any columns flagged in treat_params
  treat_cols <- names(ds$treat_params)[unlist(ds$treat_params)]
  
  # any columns flagged in temp_params
  temp_cols <- names(ds$temp_params)[unlist(ds$temp_params)]
  
  # union of both
  keep_cols <- union(union(sel_cols, treat_cols), temp_cols)

  out <- ds$data[, keep_cols, drop = FALSE]
  
  # rename treat cols with prefix, if its CO2 duplicate the column so we don't loose it
  treat_detected <- names(out)[names(out) %in% treat_cols]
  co2_pattern <- "^CO2$|^CO2_2$|^CO2_3$"
  for (nm in treat_detected) {
    if (grepl(co2_pattern, nm)) {
      # duplicate column with prefixed name
      out[[paste0("DETECTED_TREAT_", nm)]] <- out[[nm]]
    } else {
      # simple rename
      names(out)[names(out) == nm] <- paste0("DETECTED_TREAT_", nm)
    }
  }
  
  # add FINAL_TREAT if treatment columns exist
  detected_cols <- names(out)[startsWith(names(out), "DETECTED_TREAT_")]
  if (length(detected_cols) > 0) {
    out <- out %>%
      unite("FINAL_TREAT", all_of(detected_cols),
            sep = "-", remove = FALSE, na.rm = FALSE)
    
    # drop detected treat columns after creating FINAL_TREAT
    out <- out[, !names(out) %in% detected_cols, drop = FALSE]
  } else {
    out$FINAL_TREAT <- rep(NA_character_, nrow(out))
  }
  
  # split the data frame into sub-data frames by `sel_cols_comment`
  split_cols <- intersect(sel_cols, names(out))
  
  if (length(split_cols) == 0) {
    nested <- list(NO_MATCH = out)
  } else {
    grouped <- split(split_cols, sel_cols_comment[split_cols])
    
    # make nested list of data frames
    nested <- map(grouped, function(cols) {
      out[, c(cols, setdiff(names(out), sel_cols)), drop = FALSE]
    })
  }
  
  nested
}) |> set_names(map_chr(datasets, "doi"))



# Recursive function to rename columns in nested lists
drop_numeric_suffix <- function(x) {
  if (is.data.frame(x)) {
    names(x) <- str_remove(names(x), "_\\d+$")  # remove numeric suffix
    return(x)
  } else if (is.list(x)) {
    return(map(x, drop_numeric_suffix))          # recurse into sublists
  } else {
    return(x)  # leave other objects untouched
  }
}

# Apply to carb_only
carb_only <- drop_numeric_suffix(carb_only)

# Recursive function to check duplicates in nested lists
check_duplicate_cols <- function(x, path = character()) {
  if (is.data.frame(x)) {
    dup_cols <- names(x)[duplicated(names(x))]
    if (length(dup_cols) > 0) {
      return(list(path = path, duplicates = dup_cols))
    } else {
      return(NULL)
    }
  } else if (is.list(x)) {
    # recurse into sublists
    res <- map2(x, names(x), ~check_duplicate_cols(.x, c(path, .y)))
    return(compact(res))
  } else {
    return(NULL)
  }
}

# Apply to carb_only
duplicates_found <- check_duplicate_cols(carb_only)
dups <- map_depth(duplicates_found,.depth = 2, ~ .x[["duplicates"]])
names(duplicates_found)

# It seems like they didn't carry over treatment day to the seacarb columns for these datasets. Or there is extra Temperature data. Fixing the temperature one, dropping the carbonate ones:

carb_only <- map_depth(carb_only,.depth = 2, function(x) {
  nm <- names(x) # get names
  temp_idx <- which(nm == "Temp") #index that are "Temp"
  if (length(temp_idx) > 1) {
    nm[temp_idx[-1]] <- paste0("Temp", seq_along(temp_idx)[-1])
  }
  names(x) <- nm
  x
})

# Check duplicates again
duplicates_found <- check_duplicate_cols(carb_only)
dups <- map_depth(duplicates_found,.depth = 2, ~ .x[["duplicates"]])
names(duplicates_found)

# only three studies now - remove
carb_only <- carb_only[!names(carb_only) %in% names(duplicates_found)]

# combine time vars
carb_merged <- map(carb_only,~ bind_rows(.x,.id = "TIME_TREAT")) %>% bind_rows(.id = "Study_DOI")
carb_merged <- carb_merged %>% mutate(DIC_calc = CO2 + `[HCO3]-` +`[CO3]2-`)

# Duplicate data for multiple Temperature conditions
carb_merged <- carb_merged %>% pivot_longer(c(Temp,Temp2,Temp3),names_to = "temp_replication",values_to = "Temp",values_drop_na = T)

# recalc total pH with correct parameters
carb_merged[!is.na(carb_merged$Temp),"pH_recalc"] <- carb_merged %>% subset(!is.na(Temp)) %$% seacarb::carb(flag = 10,var1 = `[HCO3]-`/1e6, `[CO3]2-`/1e6,T = Temp, pHscale = "T") %>% pull(pH)

# treat multiple reported temperatures as seperate treatments
carb_merged <- carb_merged %>%
  mutate(FINAL_TREAT = if_else(temp_replication %in% c("Temp2","Temp3"),
                               true = paste(FINAL_TREAT,temp_replication,sep = ", "),
                               false = FINAL_TREAT))

carb_merged_sum <- carb_merged %>%
  group_by(FINAL_TREAT,TIME_TREAT,Study_DOI) %>% # leave out time treat to average in
  summarise(CO2_mean = mean(CO2, na.rm = T),
            HCO3_mean = mean(`[HCO3]-`, na.rm = T),
            CO3_mean = mean(`[CO3]2-`, na.rm = T),
            Temp_mean = mean(Temp, na.rm = T),
            DIC_mean = mean(DIC_calc, na.rm = T),
            pHT_mean = mean(pH_recalc,na.rm = T),
            
            Temp_sd = sd(Temp, na.rm = T),
            CO2_sd = sd(CO2, na.rm = T),
            CO3_sd = sd(`[CO3]2-`, na.rm = T),
            HCO3_sd = sd(`[HCO3]-`, na.rm = T),
            DIC_sd = sd(DIC_calc, na.rm = T),
            pHT_sd = sd(pH_recalc,na.rm = T)
            ) %>% ungroup()

OA_ICC <- as.data.frame(carb_merged_sum)

# last screen step - remove treatments without CO2_mean, HCO3_mean, CO3_mean, or pHT_mean
length(unique(OA_ICC$Study_DOI))
nrow(OA_ICC)

OA_ICC <- OA_ICC %>% subset(!(is.na(CO2_mean) | is.na(HCO3_mean) | is.na(CO3_mean) | is.na(pHT_mean)))

length(unique(OA_ICC$Study_DOI))
nrow(OA_ICC)

# clean up
rm(list = setdiff(ls(), c("datasets","OA_ICC")))

# write
write.csv(OA_ICC,here("data/OA_ICC_Data/OA_ICC_data_merged.csv"))

# list final stats
summary(OA_ICC %>% select(FINAL_TREAT,TIME_TREAT,Study_DOI,CO2_mean,
                          HCO3_mean,CO3_mean,Temp_mean,DIC_mean,pHT_mean))



```

Recreate OA-ICC Portal Information Figure
```{r}
### information here was pulled directly from OA-ICC website for panels A and B
years <- selection %>% ggplot() + geom_histogram(aes(Year),binwidth = 1) +
  theme_minimal() + ylab("# of Studies") + scale_x_continuous(breaks = c(2000,2005,2010,2015,2020,2025))

Incubation_type <- tibble(
  Container = factor(c("Containers (<20L)","Containers (20-1000L)","Mesocosm","No Data")),
  `Number of Studies` = c(205,5,1,2)
  ) %>%
  mutate(Container = forcats::fct_rev(Container)) %>%
  ggplot(aes(`Number of Studies`,Container)) +
  geom_col() + labs(y = "") + theme_minimal()

sp_type <- tibble(
  Container = factor(c("Haptophyta","Ochrophyta","Cyanobacteria","Myzozoa",
                       "Chlorophyta","Bacillariophyta","Arthropoda","Charophyta","Proteobacteria")),
  `Number of Studies` = c(93,62,28,24,12,5,3,1,1)
  ) %>%
  mutate(Container = forcats::fct_rev(Container)) %>%
  ggplot(aes(`Number of Studies`,Container)) +
  geom_col() + labs(y = "") + theme_minimal()

design <- c("AC\nBC")
 Incubation_type + sp_type + years +
   patchwork::plot_layout(design = design, ncol = 2,heights = c(0.5,1),widths = c(0.5,1),axis_titles = "collect") +
   patchwork::plot_annotation(tag_levels = "a")

```