---
title: "OA-ICC Biome Condition Assignment"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

This data compares annual biome averages and OA-ICC treatments to assess which biome average the data is closest too. It also provide some summary. Final data.frame output matches OA_ICC_data_final.csv. 

Load Packages and set working directory. 
```{r,warning=F}
# needed packages - install if necessary 

library(tidyverse) # data formatting
library(ggplot2) # plotting 
library(concaveman) # calc hulls
library(sf)
library(here) # set wd at cloned repo
library(patchwork) #plotting

biomes <- read.csv(here("data/product_data_processed/wide_annual_biomes_values.csv")) %>% as_tibble() # load annual biome data REFACTOR THIS
OA_ICC <-  read.csv(here("data/OA_ICC_Data/OA_ICC_data_merged.csv")) # load qc-ed OA-ICC data
OA_ICC <- as.data.frame(OA_ICC) # insure correct class 

# no ice biomes
#noice <- delta %>% subset(type != "ICE") %>% subset(!(scenario == "ssp585" & period %in% c("1850s","2000s")))

```

Helper Functions
```{r}
# helper function to determine 2 sd range
within_k_sd <- function(x, mean, sd, k = 2) {
  (x > (mean - (k * sd))) & (x < (mean + (k * sd)))
}

# generate 2 sd regions for each pair of vars 
generate_carbon_regions <- function(sd_level = 2, co2, co2sd, hco3, hco3sd, co3, co3sd, pH, pHsd, temp, tempsd) {
  # set sd level
  co2sd <- co2sd * sd_level
  hco3sd <- hco3sd * sd_level
  co3sd <- co3sd * sd_level
  pHsd <- pHsd * sd_level
  tempsd <- tempsd * sd_level

  df <- data.frame(co2, co2sd, hco3, hco3sd, co3, co3sd, pH, pHsd, temp, tempsd)

  pairs <- combn(c("co2", "hco3", "co3", "pH", "temp"), 2, simplify = T) %>%
    t() %>%
    data.frame()

  # get sd clouds
  expand_points <- function(df, n_angles = 16) {
    angles <- seq(0, 2 * pi, length.out = n_angles + 1)[-(n_angles + 1)]

    expanded <- lapply(1:nrow(df), function(i) {
      data.frame(
        x = df$x[i] + df$sx[i] * cos(angles),
        y = df$y[i] + df$sy[i] * sin(angles)
      )
    })
    bind_rows(expanded)
  }

  # get regions
  apply(pairs, 1, function(entry) {
    orig_df <- data.frame(
      x = df[, entry[1]],
      y = df[, entry[2]],
      sx = df[, paste0(entry[1], "sd")],
      sy = df[, paste0(entry[2], "sd")]
    )
    expanded_pts <- orig_df %>% expand_points()

    # use concaveman
    polygon_pts <- concaveman::concaveman(
      as.matrix(expanded_pts),
      concavity = 1000, # 1000 is close to convex for this data
      length_threshold = 0
    )

    # return
    polygon_df <- as.data.frame(polygon_pts)
    colnames(polygon_df) <- c(entry[1], entry[2])
    list(polygon_df, orig_df)
  })
}

# plot outcomes 
plot_carbon_region <- function(region_list, oa_icc = T, oa_df = OA_ICC, fill = "grey50", color = "black") {
  labs <- list(
    hco3 = expression(paste("[HCO"[3]^"-", "]", " (", "umol ", kg^{-1}, ")")),
    co2  = expression("[CO"[2] * "] (umol kg"^{-1} * ")"),
    co3  = expression(paste("[CO"[3]^"-2", "]", " (", "umol ", kg^{-1}, ")")),
    temp = "Temperature (°C)",
    pco2 = expression("pCO"[2] * ""),
    pH   = "pH (Total Scale)"
  )
  
  oa_icc_key <- data.frame(
    row.names = c("co2", "hco3", "co3", "pH", "temp"),
    icc = c("CO2_mean", "HCO3_mean", "CO3_mean", "pHT_mean", "Temp_mean")
  )
  
  sqrt_vars <- c("co2", "hco3", "co3")  # variables that get log10 transform
  
  make_scale <- function(var, axis) { # make scales 
    scale_fn <- if (axis == "x") scale_x_log10 else scale_y_log10
    if (var %in% sqrt_vars) scale_fn() else NULL
  }
  
  plots <- lapply(seq_along(region_list), function(i) {
    polygon_df <- region_list[[i]][[1]]
    raw_pts    <- region_list[[i]][[2]]
    col_names  <- colnames(polygon_df)
    
    if (oa_icc) {
      ggplot() +
        geom_point(
          aes(x = .data[[oa_icc_key[col_names[1], "icc"]]],
              y = .data[[oa_icc_key[col_names[2], "icc"]]]),
          color = "grey30", alpha = 0.3, shape = 19,
          data = oa_df
        ) +
        geom_polygon(
          data = polygon_df,
          aes(x = .data[[col_names[1]]], y = .data[[col_names[2]]]),
          fill = fill, alpha = 0.3, color = color, linewidth = 1,
          inherit.aes = FALSE
        ) +
        theme_minimal() +
        labs(x = labs[[col_names[1]]], y = labs[[col_names[2]]]) +
        make_scale(col_names[1], "x") +
        make_scale(col_names[2], "y")
    } else {
      ggplot() +
        geom_polygon(
          data = polygon_df,
          aes(x = .data[[col_names[1]]], y = .data[[col_names[2]]]),
          fill = fill, alpha = 0.3, color = color, linewidth = 1,
          inherit.aes = FALSE
        ) +
        geom_errorbar(aes(x = x, y = y, ymin = y - sy, ymax = y + sy),
                      width = 0, color = "gray40", data = raw_pts) +
        geom_errorbarh(aes(x = x, y = y, xmin = x - sx, xmax = x + sx),
                       height = 0, color = "gray40", data = raw_pts) +
        geom_point(size = 2, aes(x = x, y = y), data = raw_pts) +
        theme_minimal() +
        labs(x = labs[[col_names[1]]], y = labs[[col_names[2]]])
    }
  })
  plots
}

# check OA-ICC and lab points 
check_points <- function(co2, hco3, co3, pH, temp, carbon_regions) {
  df <- data.frame(co2, hco3, co3, pH, temp)

  results <- map(carbon_regions, function(region) {
    pair <- colnames(region[[1]])
    test_points <- df[pair]

    # Create polygon boundary
    poly_sf <- st_polygon(list(as.matrix(region[[1]])))

    # Convert to sf objects 
    points_sf <- st_as_sf(test_points, coords = c(pair))
    poly_sfc <- st_sfc(poly_sf)

    # Test containment
    inside <- st_within(points_sf, poly_sfc, sparse = FALSE)
    inside <- as.vector(inside)
    inside <- list(inside)
    names(inside) <- paste(pair, collapse = "_")
    inside
  }) %>%
    map(bind_cols) %>%
    bind_cols()

  results$ocean_relevant <- apply(results, 1, all)

  bind_cols(df, results)
}

```

Calculate Biome Multi-driver Regions
```{r}

carbon_regions <- generate_carbon_regions(
  co2 = biomes$areaW_mean_abs_aqueous,
  co2sd = biomes$sd_abs_aqueous,
  hco3 = biomes$areaW_mean_abs_bicarbonate,
  hco3sd = biomes$sd_abs_bicarbonate,
  co3 = biomes$areaW_mean_abs_carbonate,
  co3sd = biomes$sd_abs_carbonate,
  pH = biomes$areaW_mean_abs_pH,
  pHsd = biomes$sd_abs_pH,
  temp = biomes$areaW_mean_abs_tos,
  tempsd = biomes$sd_abs_tos
)

```

Plot Multidriver Regions Supp. Figures
```{r fig.height=6, fig.width=8, warning=FALSE}

plots <- plot_carbon_region(carbon_regions,oa_icc = F)

wrap_plots(
  plots[[1]], plot_spacer(), plot_spacer(), plot_spacer(),
  plots[[2]], plots[[5]], plot_spacer(), plot_spacer(),
  plots[[3]], plots[[6]], plots[[8]], plot_spacer(),
  plots[[4]], plots[[7]], plots[[9]], plots[[10]],
  ncol = 4, nrow = 4
) + plot_layout(axes = "collect")

plots <- plot_carbon_region(carbon_regions,color = "red",fill = NA,
                            oa_df = OA_ICC)

wrap_plots(
  plots[[1]], plot_spacer(), plot_spacer(), plot_spacer(),
  plots[[2]], plots[[5]], plot_spacer(), plot_spacer(),
  plots[[3]], plots[[6]], plots[[8]], plot_spacer(),
  plots[[4]], plots[[7]], plots[[9]], plots[[10]],
  ncol = 4, nrow = 4
) + plot_layout(axes = "collect")


```

Assign to treatments 
```{r}

# check oa-icc points against multidriver regions, check if relavant 
results <- check_points(
  co2 = OA_ICC$CO2_mean,
  hco3 = OA_ICC$HCO3_mean,
  co3 = OA_ICC$CO3_mean,
  pH = OA_ICC$pHT_mean,
  temp = OA_ICC$Temp_mean,
  carbon_regions = carbon_regions
)

OA_ICC$ocean_relevant <- results$ocean_relevant

# calc high/low limits for later
carblims <- data.frame(
  co2_lower = min(biomes$areaW_mean_abs_aqueous - biomes$sd_abs_aqueous * 2),
  co2_upper = max(biomes$areaW_mean_abs_aqueous + biomes$sd_abs_aqueous * 2),
  hco3_lower = min(biomes$areaW_mean_abs_bicarbonate - biomes$sd_abs_bicarbonate * 2),
  hco3_upper = max(biomes$areaW_mean_abs_bicarbonate + biomes$sd_abs_bicarbonate * 2),
  co3_lower = min(biomes$areaW_mean_abs_carbonate - biomes$sd_abs_carbonate * 2),
  co3_upper = max(biomes$areaW_mean_abs_carbonate + biomes$sd_abs_carbonate * 2),
  ph_lower = min(biomes$areaW_mean_abs_pH - biomes$sd_abs_pH * 2),
  ph_upper = max(biomes$areaW_mean_abs_pH + biomes$sd_abs_pH * 2),
  temp_upper = max(biomes$areaW_mean_abs_tos + biomes$sd_abs_tos * 2),
  temp_lower = min(biomes$areaW_mean_abs_tos - biomes$sd_abs_tos * 2)
)

# assign to treatments 
oa_icc_hits <- list()
for (i in 1:nrow(OA_ICC)) {
  # ind hits
  co2_hits <- within_k_sd(OA_ICC[i, "CO2_mean"], biomes$areaW_mean_abs_aqueous, biomes$sd_abs_aqueous)
  hco3_hits <- within_k_sd(OA_ICC[i, "HCO3_mean"], biomes$areaW_mean_abs_bicarbonate, biomes$sd_abs_bicarbonate)
  co3_hits <- within_k_sd(OA_ICC[i, "CO3_mean"], biomes$areaW_mean_abs_carbonate, biomes$sd_abs_carbonate)
  temp_hits <- within_k_sd(OA_ICC[i, "Temp_mean"], biomes$areaW_mean_abs_tos, biomes$sd_abs_tos)
  ph_hits <- within_k_sd(OA_ICC[i, "pHT_mean"], biomes$areaW_mean_abs_pH, biomes$sd_abs_pH)

  cols <- c("scenario", "biome", "loc", "type", "period") # to assign
  if (OA_ICC[i, "ocean_relevant"]) {
    # what point is it closest too - evaluate all points so there is no sd bias
    sd_vec <- c(
      sd(biomes$areaW_mean_abs_aqueous),
      sd(biomes$areaW_mean_abs_bicarbonate),
      sd(biomes$areaW_mean_abs_carbonate),
      sd(biomes$areaW_mean_abs_tos),
      sd(biomes$areaW_mean_abs_pH)
    )

    d <- sqrt(
      ((OA_ICC[i, "CO2_mean"] - biomes$areaW_mean_abs_aqueous) / sd_vec[1])^2 +
        ((OA_ICC[i, "HCO3_mean"] - biomes$areaW_mean_abs_bicarbonate) / sd_vec[2])^2 +
        ((OA_ICC[i, "CO3_mean"] - biomes$areaW_mean_abs_carbonate) / sd_vec[3])^2 +
        ((OA_ICC[i, "Temp_mean"] - biomes$areaW_mean_abs_tos) / sd_vec[4])^2 +
        ((OA_ICC[i, "pHT_mean"] - biomes$areaW_mean_abs_pH) / sd_vec[5])^2
    )

    closest_idx <- which.min(d)

    oa_icc_hits[[i]] <- biomes[closest_idx, ]
    OA_ICC[i, cols] <- biomes[closest_idx, cols]
  } else {
    oa_icc_hits[[i]] <- NULL
    OA_ICC[i, cols] <- "Non-relevant"

    OA_ICC[i, "miss_co2_dir"] <- ifelse(OA_ICC[i, "CO2_mean"] > carblims$co2_upper, "high",
      ifelse(OA_ICC[i, "CO2_mean"] < carblims$co2_lower, "low", NA)
    )
    OA_ICC[i, "miss_hco3_dir"] <- ifelse(OA_ICC[i, "HCO3_mean"] > carblims$hco3_upper, "high",
      ifelse(OA_ICC[i, "HCO3_mean"] < carblims$hco3_lower, "low", NA)
    )
    OA_ICC[i, "miss_co3_dir"] <- ifelse(OA_ICC[i, "CO3_mean"] > carblims$co3_upper, "high",
      ifelse(OA_ICC[i, "CO3_mean"] < carblims$co3_lower, "low", NA)
    )
    OA_ICC[i, "miss_pH_dir"] <- ifelse(OA_ICC[i, "pHT_mean"] > carblims$ph_upper, "high",
      ifelse(OA_ICC[i, "pHT_mean"] < carblims$ph_lower, "low", NA)
    )
    OA_ICC[i, "miss_temp_dir"] <- ifelse(OA_ICC[i, "Temp_mean"] > carblims$temp_upper, "high",
      ifelse(OA_ICC[i, "Temp_mean"] < carblims$temp_lower, "low", NA)
    )

    #each ion has hits, but never on the same biome (no co-occurrence)
    OA_ICC[i, "miss_cooccurrence"] <- any(co2_hits, na.rm = TRUE) &
      any(hco3_hits, na.rm = TRUE) &
      any(co3_hits, na.rm = TRUE) &
      any(temp_hits, na.rm = TRUE) &
      any(ph_hits, na.rm = TRUE)
  }
}

# for non aligned treatments, get directionality of carbonate relative to ocean (i.e. high or low)

# check high
OA_ICC$miss_high <- (is.na(OA_ICC$miss_co2_dir) | OA_ICC$miss_co2_dir == "high") &
  (is.na(OA_ICC$miss_hco3_dir) | OA_ICC$miss_hco3_dir == "high") &
  (is.na(OA_ICC$miss_pH_dir) | OA_ICC$miss_pH_dir == "low") &
  !(is.na(OA_ICC$miss_co2_dir) & is.na(OA_ICC$miss_hco3_dir) & is.na(OA_ICC$miss_co3_dir) & is.na(OA_ICC$miss_pH_dir))

# check low
OA_ICC$miss_low <- (is.na(OA_ICC$miss_co2_dir) | OA_ICC$miss_co2_dir == "low") &
  (is.na(OA_ICC$miss_hco3_dir) | OA_ICC$miss_hco3_dir == "low") &
  (is.na(OA_ICC$miss_pH_dir) | OA_ICC$miss_pH_dir == "high") &
  !(is.na(OA_ICC$miss_co2_dir) & is.na(OA_ICC$miss_hco3_dir) & is.na(OA_ICC$miss_co3_dir) & is.na(OA_ICC$miss_pH_dir))

# combine
OA_ICC <- OA_ICC %>%
  rowwise() %>%
  mutate(
    dir = (if_else(miss_high, "high", NA)),
    dir = (if_else(miss_low, "low", dir))
  )

# things that were not high or low are missaligned
OA_ICC[which(OA_ICC$ocean_relevant == F & is.na(OA_ICC$dir) & OA_ICC$miss_cooccurrence != T), "miss_cooccurrence"] <- T

OA_ICC %>%
  count(dir, ocean_relevant, miss_cooccurrence, name = "n") %>%
  arrange(desc(n))

write.csv(OA_ICC,here("data/OA_ICC_Data/OA_ICC_data_final.csv"))

```

Plotting Supp. Figures and Stats for Paper
```{r fig.height=8, fig.width=8}
# label formatting 
dhco3 <- expression(paste("Δ [HCO"[3]^"-","]", " (", "umol ", kg^{-1}, ")"))
dco2 <- expression("Δ [CO"[2]*"] (umol kg"^{-1}*")")
dco3 <- expression(paste("Δ [CO"[3]^"-2","]", " (", "umol ", kg^{-1}, ")"))
dtemp <- "Δ Temperature (°C)"
dpco2 <- expression("Δ pCO"[2]*"")

hco3 <- expression(paste("Δ [HCO"[3]^"-","]", " (", "umol ", kg^{-1}, ")"))
co2 <- expression("Δ [CO"[2]*"] (umol kg"^{-1}*")")
co3 <- expression(paste("Δ [CO"[3]^"-2","]", " (", "umol ", kg^{-1}, ")"))
temp <- "Δ Temperature (°C)"
pco2 <- expression("Δ pCO"[2]*"")
pH <- "pH (Total Scale)"

# pie plotting
# combine for stats on a study level
rel_plot_sum <- OA_ICC %>%
  group_by(Study_DOI) %>%
  summarise(
    any_relevant = any(ocean_relevant),
    any_not_relevant = any(!ocean_relevant),
    all_relevant = all(ocean_relevant),
    all_not_relevant = all(!ocean_relevant),
    mixed_relevant = any(ocean_relevant) & any(!ocean_relevant),
    has_1850s = any(period == "1850s"),
    has_2000s = any(period == "2000s"),
    has_2090s = any(period == "2090s"),
    has_ssp245 = any(period == "2090s" & scenario == "ssp245"),
    has_ssp585 = any(period == "2090s" & scenario == "ssp585"),
    has_high = any(dir == "high"),
    has_low = any(dir == "low"),
    has_miss = any(miss_cooccurrence),
    has_EQU = any(type == "EQU"),
    has_STPS = any(type == "STPS"),
    has_STSS = any(type == "STSS"),
    has_SPSS = any(type == "SPSS"),
    .groups = "drop"
  )

sum(OA_ICC$ocean_relevant) # 1402 biome relavent treatments 
sum(OA_ICC$ocean_relevant)/nrow(OA_ICC) # 57% of treatments biome relavent
rel_plot_sum %>% summarise(across(-Study_DOI, ~ sum(.x) / 196)) # more stats - relative
rel_plot_sum %>% summarise(across(-Study_DOI, ~ sum(.x))) # more stats - # of studies 

# supp figure 
OA_ICC <- OA_ICC %>%
  mutate(category = case_when(
    miss_cooccurrence      ~ "Misaligned",
    dir == "high"          ~ "Higher Inorganic Carbon",
    dir == "low"           ~ "Lower Inorganic Carbon",
    ocean_relevant         ~ "Biome Relevant"
  ))

ggplot(OA_ICC, aes(y= pHT_mean,x = CO2_mean , color = category)) +
  geom_point(shape = 19,alpha = 0.3) +
  scale_color_manual(values = c(
    "Background"     = "grey80",
    "Higher Inorganic Carbon"        = "#E05C3A",
    "Lower Inorganic Carbon"         = "#4A90D9",
    "Misaligned"     = "#9B59B6",
    "Biome Relevant" = "#F5C518"
  )) +
  theme_minimal() +
  lims(x = c(0, 100)) +
  labs(x = co2,y = pH,color = "OA-ICC Treatment Alignment") +
  plot_spacer() +
ggplot(OA_ICC, aes(CO2_mean, HCO3_mean, color = category)) +
  geom_point(shape = 19,alpha = 0.3) +
  scale_color_manual(values = c(
    "Background"     = "grey80",
    "Higher Inorganic Carbon"        = "#E05C3A",
    "Lower Inorganic Carbon"         = "#4A90D9",
    "Misaligned"     = "#9B59B6",
    "Biome Relevant" = "#F5C518"
  )) +
  theme_minimal() +
  lims(x = c(0, 100), y = c(0, 3000)) +
  labs(x = co2,y = hco3,color = "OA-ICC Treatment Alignment") +
  ggplot(OA_ICC, aes(y= HCO3_mean,x = pHT_mean , color = category)) +
  geom_point(shape = 19,alpha = 0.3) +
  scale_color_manual(values = c(
    "Background"     = "grey80",
    "Higher Inorganic Carbon"        = "#E05C3A",
    "Lower Inorganic Carbon"         = "#4A90D9",
    "Misaligned"     = "#9B59B6",
    "Biome Relevant" = "#F5C518"
  )) +
  theme_minimal() +
  lims(y = c(0, 3000)) +
  labs(x = pH,y = hco3,color = "OA-ICC Treatment Alignment") +
  plot_layout(ncol = 2,nrow = 2,guides = "collect",axes = "collect",) 


rel_plot_sum$has_low[is.na(rel_plot_sum$has_low)] <- FALSE
rel_plot_sum$has_high[is.na(rel_plot_sum$has_high)] <- FALSE
rel_plot_sum$has_miss[is.na(rel_plot_sum$has_miss)] <- FALSE

# pie plot
pie1_a <- data.frame(
  `all rel` = sum(rel_plot_sum$all_relevant),
  `no rel` = sum(rel_plot_sum$all_not_relevant),
  `mixed` = sum(rel_plot_sum$mixed_relevant)
) %>%
  pivot_longer(everything()) %>%
  ggplot(aes(x = "", y = value, fill = name)) +
  geom_bar(stat = "identity", width = 1, color = "black", linewidth = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  labs(fill = "") +
  scale_fill_manual(values = c("white", "grey50", "black")) +
  geom_label(aes(label = value), color = "black", fill = "white", position = position_stack(vjust = 0.5))

b <- ggVennDiagram::ggVennDiagram(list(
  `1850s` = rel_plot_sum$Study_DOI[rel_plot_sum$has_1850s],
  `2000s` = rel_plot_sum$Study_DOI[rel_plot_sum$has_2000s],
  `2090s` = rel_plot_sum$Study_DOI[rel_plot_sum$has_2090s]
), label = "count") +
  scale_fill_gradient(high = "#0096FF", low = "white", name = "# of Studies", limits = c(0, 90))

c <- ggVennDiagram::ggVennDiagram(list(
  `SSP2-4.5` = rel_plot_sum$Study_DOI[rel_plot_sum$has_ssp245],
  `SSP5-8.5` = rel_plot_sum$Study_DOI[rel_plot_sum$has_ssp585]
), label = "count") + scale_fill_gradient(high = "#0096FF", low = "white", name = "# of Studies", limits = c(0, 90))

d <- ggVennDiagram::ggVennDiagram(list(
  `High` = rel_plot_sum$Study_DOI[rel_plot_sum$has_high],
  `Low` = rel_plot_sum$Study_DOI[rel_plot_sum$has_low],
  # `Rel` = rel_plot_sum$Study_DOI[rel_plot_sum$any_relevant]#,
  `miss` = rel_plot_sum$Study_DOI[rel_plot_sum$has_miss]
), label = "count") + scale_fill_gradient(high = "#0096FF", low = "white", name = "# of Studies", limits = c(0, 90))

e <- ggVennDiagram::ggVennDiagram(list(
  `EQU` = rel_plot_sum$Study_DOI[rel_plot_sum$has_EQU],
  `STPS` = rel_plot_sum$Study_DOI[rel_plot_sum$has_STPS],
  `SPSS` = rel_plot_sum$Study_DOI[rel_plot_sum$has_SPSS],
  `STSS` = rel_plot_sum$Study_DOI[rel_plot_sum$has_STSS]
), label = "count") + scale_fill_gradient(high = "#0096FF", low = "white", name = "# of Studies", limits = c(0, 90))

pie1_a + b + c + d + plot_layout(guides = "collect", nrow = 2) + plot_annotation(tag_levels = "A")

```

