---
title: "Data proccessing for Holm et al. - choosing phytoplankton experimental evolution conditions with climate models"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

### Data Analysis Section ###

Load Data and Packages
```{r}
# load packages
library(tidyverse) # data processing, plotting
library(tidyterra) # spatial data processing
library(terra) # spatial data processing
library(rnaturalearth) # spatial data processing
library(purrr) # Functional Programming
library(tidyr) # data processing
library(seacarb) # carbonate calculations
library(errors) # error propagation
library(patchwork) # plotting
library(ggplot2) # plotting
library(magrittr) # data processing
library(furrr) # parallel
library(data.table) # data processing

theme_set(theme_minimal())

base_loc <- "C:/Users/henryholm/OneDrive - Woods Hole Oceanographic Institution/Desktop/Galen_CMIP_Collab/raw_data/"

#point to local intance of the ORCCparams repo
local <- "C:/Users/henryholm/Documents/GitHub/"

# contains data on culture pH at variable cell densities
pHcd_all<- read.csv(paste0(local,"ORCCparams/lab/data/all_pH_training.csv"))

# Biome data
biomes <- terra::rast(paste0(base_loc, "Time_Varying_Biomes.nc"))

# Format Galen biome data
#biomes <- flip(t(biomes))
crs(biomes) <- "+proj=longlat +ellips=WGS84"
biomes_vec <- as.polygons(biomes)
biomes_vec$MeanBiomes <- c(
  "NP_ICE", "NP_SPSS", "NP_STSS", "NP_STPS",
  "Pac_EQU_W", "Pac_EQU_E", "SP_STPS", "NA_ICE",
  "NA_SPSS", "NA_STSS", "NA_STPS", "Atl_EQU",
  "SA_STPS", "IND_STPS", "SO_STSS", "SO_SPSS",
  "SO_ICE"
)

# weight data
weights <- terra::rast(paste0(base_loc, "area_weights_km2.nc"))

w_df <- weights %>% as.data.frame(xy = T) %>%
  pivot_longer(
    cols = -c(x, y),
    values_to = "area_km2"
  ) %>% select(-name) %>% as.data.table()


# GLODAP Mapped Data
loc <- paste0(base_loc, "GLODAPv2.2016b_MappedClimatologies/")

# file names
glodap_map_files <- list(
  temp = "GLODAPv2.2016b.temperature.nc",
  TCO2 = "GLODAPv2.2016b.TCO2.nc",
  pH = "GLODAPv2.2016b.pHtsinsitutp.nc",
  TAlk = "GLODAPv2.2016b.TAlk.nc",
  salinity = "GLODAPv2.2016b.salinity.nc"
)


# load files and add lat long projection
glodap_map <- list()
for (file in names(glodap_map_files)) {
  glodap_map[[file]] <- terra::rast(paste0(loc, glodap_map_files[[file]]))
  crs(glodap_map[[file]]) <- "+proj=longlat +ellips=WGS84"
}

# Future Mapped Data - CMIP6
plan(multisession, workers = parallel::detectCores()) # parallel loading

# Set base directory and list NetCDF files
base_dir <- "C:/Users/henryholm/Documents/IPCC_Galen_Data/CMIP-6/decade_averages/monthly/"
all_files <- list.files(base_dir, full.names = TRUE, recursive = TRUE, pattern = "\\.nc$")
all_files <- all_files[!grepl("ensemble", all_files, ignore.case = TRUE)]

# Extract metadata from filenames
file_info <- tibble::tibble(
  file = all_files,
  model = str_extract(basename(all_files), "(?<=timeseries_)[^_]+"),
  scenario = str_extract(file, "ssp[0-9]+"),
  variable = str_extract(basename(all_files), "^[^_]+")
)

# Initialize output list
cmip6 <- list()

for (i in 1:nrow(file_info)) {
  cmip6[[file_info[[i, "model"]]]][[file_info[[i, "scenario"]]]][[file_info[[i, "variable"]]]] <- file_info[[i, "file"]]
}

cmip6 <- map_depth(cmip6, 3, terra::rast)

# BATS data
loc <- "C:/Users/henryholm/Documents/GitHub/ORCCparams/climate_models/data/bats_bottle_formatted.txt"
BATS <- read.delim(loc, header = T)

# max months for 'bloom' - get formal analysis from Anna

bloom_months <- tibble::tibble(
  biome = c(
    "NP_SPSS",
    "NP_STSS",
    "NA_SPSS",
    "NA_STSS",
    "SO_STSS",
    "SO_SPSS"
  ),
  month = list(
    c(5, 6, 7),
    c(4, 5, 6),
    c(6, 7, 8),
    c(4, 5, 6),
    c(12, 1, 2),
    c(12, 1, 2)
  )
)

# Thomas Data
base_loc <- "C:/Users/henryholm/OneDrive - Woods Hole Oceanographic Institution/Desktop/Galen_CMIP_Collab/raw_data/"
thomas <- list()
thomas$max <- read.csv(paste0(base_loc,"thomas_2012/thomas.sm.table.s5.csv"))
thomas$loc <- read.csv(paste0(base_loc,"thomas_2012/thomas.sm.table.s6.csv"))

# Remove uneeded temp objects
remove(file, i, loc, all_files,base_dir,base_loc,file_info)

```

PyCO2Sys Calculations - Set up Python Env
```{r}
library(reticulate)

virtualenv_create("co2sys-env",python = "C:/Users/henryholm/AppData/Local/Python/pythoncore-3.14-64/python.exe")
use_virtualenv("co2sys-env", required = TRUE)
virtualenv_install("co2sys-env", "PyCO2SYS")

pyco2 <- import("PyCO2SYS")
```

GLODAP Data Formatting
This chunk of code separates the mapped GLODAP surface data into Biome regions. Additionally it uses seacarb to calculate the full carbonate system using DIC, Alk, T, and S from observation data.
```{r}

# fix lat long orientation
glodap_map <- purrr::map(glodap_map, .progress = T, function(x) {
  slice <- crop(x, ext(360, 380, -90, 90)) # crop the 20 degree section above 360
  ext(slice) <- ext(0, 20, -90, 90) # reassign it as 0 to 20
  out <- terra::merge(slice, x) # merge it back with the full map
  out <- crop(out, ext(0, 360, -90, 90)) # crop the excess again
  out <- terra::rotate(out) # reorient map with 360 as 0
  return(out)
})

# plotting# plotting# plotting
ggplot() +
  geom_spatraster(data = glodap_map[["temp"]][[1]]) +
  geom_spatvector(data = biomes_vec,fill = NA,color = 'black') +
  scale_fill_viridis_c()

# cut just to surface layers for memory handling
glodap_map_biome <- purrr::map(glodap_map, .progress = T, function(x) {
    x %>%
    subset(stringr::str_detect(names(x),"=0"))
})

# region masking
glodap_map_biome <- purrr::map(glodap_map_biome, .progress = T, function(x) {
  map(biomes_vec$MeanBiomes, function(y) {
    mask(x, mask = biomes_vec[which(biomes_vec$MeanBiomes == y)])
  })
})

# names fix
for (i in 1:length(glodap_map_biome)) {
  names(glodap_map_biome[[i]]) <- biomes_vec$MeanBiomes
}

# plotting
ggplot() +
  geom_spatraster(data = glodap_map_biome[["temp"]][["NA_STPS"]][[1]]) +
  geom_spatvector(data = biomes_vec,fill = NA,color = 'black') +
  geom_point(aes(y = 32.166667,x = -64.5)) +
  scale_fill_viridis_c()

# surface masking
glodap_surf_df <- purrr::map_depth(glodap_map_biome, 2, .progress = T, function(x) {
  x %>%
    subset(stringr::str_detect(names(x),"=0")) %>%
    terra::as.data.frame(xy = TRUE) 
})

# setting error
glodap_surf_df <- purrr::map_depth(glodap_surf_df, 2, .progress = T, function(x) {
   y <- set_errors(x[,3],x[,4])
   x[,"param_error"] <- y
   x
})

# carbonate data subsetting
glodap_surf_carb <- purrr::transpose(glodap_surf_df)

glodap_surf_carb <- purrr::map(glodap_surf_carb, .progress = T, function(x) {
  z <- data.frame(
    x = x$temp$x,
    y = x$temp$y,
    pH = x$pH$`pHtsinsitutp_depth_surface=0`,
    TAlk = x$TAlk$`TAlk_depth_surface=0` / 1e6,
    TCO2 = x$TCO2$`TCO2_depth_surface=0` / 1e6,
    S = x$salinity$`salinity_depth_surface=0`,
    Temp = x$temp$`temperature_depth_surface=0`
  ) %>%
    subset(complete.cases(.))

  return(z %$%
    carb(flag = 15, var1 = TAlk, var2 = TCO2, S = S, T = Temp) %>%
    cbind(long = z$x, lat = z$y, pH_og = z$pH))
})

# take means
glodap_surf_calc <- purrr::map_depth(glodap_surf_df, 2, .progress = T, function(x) {
   y <- set_errors(x[,3],x[,4])
   data.frame(mean = mean(y,na.rm = T),median = median(y,na.rm = T))
})

# formatting
glodap_dat_carb <- tibble::enframe(glodap_surf_carb,name = "loc",value = "data") %>%
     unnest(data)
glodap_mean_dat <- tibble::enframe(glodap_surf_calc,name = "param",value = "data") %>%
     unnest(data)
glodap_mean_dat$loc <- rep(biomes_vec$MeanBiomes,length(glodap_surf_df))
glodap_mean_dat <- glodap_mean_dat %>% unnest(data) %>% pivot_wider(values_from =  c('mean','median'),names_from = "param")

# plotting
glodap_dat_carb %>% subset(!str_detect(.$loc,"_ICE")) %>%
  ggplot() + geom_density_2d(aes(T,DIC*1e6,color = loc),alpha = 0.5,contour_var = "count") +
  geom_point(aes(mean_temp,mean_TCO2,color = loc),size = 3,data = glodap_mean_dat %>% subset(!str_detect(.$loc,"_ICE"))) +
  geom_errorbar(aes(mean_temp,mean_TCO2,
                    xmin = mean_temp-errors(mean_temp),
                    xmax = mean_temp+errors(mean_temp),
                    color = loc),
                data = glodap_mean_dat %>% subset(!str_detect(.$loc,"_ICE"))) +
  geom_errorbar(aes(mean_temp,mean_TCO2,
                    ymin = mean_TCO2-errors(mean_TCO2),
                    ymax = mean_TCO2+errors(mean_TCO2),
                    color = loc),
                data = glodap_mean_dat %>% subset(!str_detect(.$loc,"_ICE"))) 

# formatting
glodap_dat_carb <- glodap_dat_carb$loc %>%
  str_split(pattern = "_",simplify = T) %>%
  data.frame() %>%
  dplyr::rename(loc = X1,type = X2) %>%
  select(loc,type) %>%
  cbind(.,glodap_dat_carb %>% dplyr::rename(biome = loc))

# remove extra objects and group data
glodap <- list(dataframe = glodap_dat_carb,
               means = glodap_mean_dat,
               maps = glodap_map_biome)

# save surface for later
glodap$temp_map_thomas <- glodap_map$temp$`temperature_depth_surface=0`

remove(glodap_surf_df,glodap_surf_carb,glodap_surf_calc,
       glodap_map,glodap_mean_dat,glodap_map_biome,glodap_dat_carb)

```

BATS Data Formatting
This chunk of code cuts 30 years of BATS data into a slice at 10 and 100 meters and takes averages.
```{r}

# set plot theme
ggplot2::theme_set(ggplot2::theme_minimal())

# id missing values as NA 
BATS[BATS == -999.000] <- NA

# refactor
BATS$DIC <- BATS$CO2
BATS$CO2 <- NULL

# calculate water pressure
BATS$hydrostatic_pressure_bar <- seacarb::d2p(BATS$Depth,lat = 32.166667)/10

# plot raw data
ggplot(BATS) + geom_point(aes(Temp,Depth)) + scale_y_reverse() 
ggplot(BATS) + geom_point(aes(Alk,Depth)) + scale_y_reverse() +
ggplot(BATS) + geom_point(aes(DIC,Depth)) + scale_y_reverse()

BATS$month <- stringr::str_remove(BATS$yyyymmdd,pattern = "^.{4}") %>%
  stringr::str_extract(pattern = "(.{2})")

# calculate co2 enviroment
sel <- BATS %>%
  select(Depth,Alk,DIC,Temp,Sal1,hydrostatic_pressure_bar) %>%
  complete.cases()

BATS_carb <- BATS %>%
  select(Depth,decy,month,Alk,DIC,Temp,Sal1,hydrostatic_pressure_bar,
         Pro,POC,Syn,Piceu,Naneu,yyyymmdd) %>%
  subset(sel)

BATS_carb1 <- BATS_carb %>% select(-Alk,-DIC) %>% cbind(.,BATS_carb %$%
          seacarb::carbb(flag = 15, var1 = Alk/1e6, var2 = DIC/1e6,
                        T = Temp, S = Sal1, 
                        P = hydrostatic_pressure_bar))


### full carb calc
BATS_carb <- pyco2$sys(
  par1      = BATS_carb$Alk,
  par2      = BATS_carb$DIC,
  par1_type = 1L,
  par2_type = 2L,
  salinity  = BATS_carb$Sal1,
  opt_pH_scale = 1,
  temperature = BATS_carb$Temp,
  pressure  = BATS_carb$hydrostatic_pressure_bar/10, #dbar
  opt_k_carbonic = 10,
  opt_k_bisulfate = 1,
  opt_total_borate =1,
  opt_k_fluoride = 1
) %>% 
  data.frame() %>% 
  rename_with(~ paste0("pyco2sys_", .)) %>%
  cbind(BATS_carb,.)

ggplot(BATS_carb) +
  geom_point(aes(pyco2sys_pCO2,Depth)) +
  scale_y_continuous(trans = scales::transform_compose(
    scales::transform_sqrt(), 
    scales::transform_reverse()
  ))

library(ggpointdensity)
BATS_carb %>%
  mutate(CO2 = pyco2sys_CO2*1e6,
         HCO3 = pyco2sys_HCO3*1e6,
         CO3 = pyco2sys_CO3*1e6,
         DIC = pyco2sys_dic*1e6,
         ALK = pyco2sys_alkalinity*1e6,
         pCO2 = pyco2sys_pCO2,
         pH = pyco2sys_pH_total) %>%
  pivot_longer(c("pCO2","pH","CO2","HCO3","CO3","DIC","ALK","Pro","POC","Syn","Piceu","Naneu","Temp"),names_to = "carb_var") %>%
  ggplot() + geom_point(aes(value,Depth)) +
  scale_y_reverse() + lims(y = c(1000,0)) +
  geom_hline(yintercept = 200,color = 'red') +
  facet_wrap(vars(carb_var),scales = "free_x")

# depth cut
BATS_dsub <- rbind(
  BATS_carb %>% subset(Depth > 95 & Depth < 105) %>% mutate(nominal_depth = 100),
  BATS_carb %>% subset(Depth > 35 & Depth < 45) %>% mutate(nominal_depth = 40),
  BATS_carb %>% subset(Depth > 7.5 & Depth < 12.5) %>% mutate(nominal_depth = 10),
  BATS_carb %>% subset(Depth < 5) %>% mutate(nominal_depth = 0)
  )

BATS_dsub %>%
  mutate(CO2 = pyco2sys_CO2*1e6, HCO3 = pyco2sys_HCO3*1e6, CO3 = pyco2sys_CO3*1e6,
         DIC = pyco2sys_dic*1e6,ALK = pyco2sys_alkalinity*1e6,pCO2 = pyco2sys_pCO2,
         pH = pyco2sys_pH_total) %>%
  pivot_longer(c("pCO2","pH","CO2","HCO3","CO3","DIC","ALK","Temp"),
               names_to = "carb_var") %>%
  ggplot() + geom_line(aes(decy,value,color = as.character(nominal_depth))) +
  theme(legend.position = "top",strip.placement = "outside") +
  labs(y = "",x = "Time (Years)",color = "Nominal Depth (m)") +
  facet_wrap(vars(carb_var),scales = "free_y",strip.position = "left",axes = 'all_x')

BATS_dsub <- BATS_dsub %>%
  mutate(CO2 = pyco2sys_CO2, HCO3 = pyco2sys_HCO3, CO3 = pyco2sys_CO3,
         DIC = pyco2sys_dic,ALK = pyco2sys_alkalinity,pCO2 = pyco2sys_pCO2,
         pH = pyco2sys_pH_total) 

t <- spline(y = BATS_dsub$CO2, x = BATS_dsub$decy)

ggplot() + geom_point(aes(x = BATS_dsub$decy,y = BATS_dsub$CO2)) + geom_line(aes(t$x,t$y))

a <- TSA::periodogram(t$y)
plot(a$spec,log10(a$freq))

# get month averages
BATS_dsub <- BATS_dsub %>%
  subset(nominal_depth %in% c(0,10,40,100)) %>%
  mutate(yyyymmdd = as.character(yyyymmdd)) %>%
  separate(yyyymmdd,into = c("year","mmdd"),sep = 4) %>%
  separate(mmdd,into = c("month","day"),sep = 2) %>%
  group_by(year,month,nominal_depth) %>%
  summarise_if(.predicate = is.numeric,.funs = ~ mean(.x, na.rm = T))

BATS_dsub$decade <- cut(BATS_dsub$decy,breaks = c(1990,2000,2010,2020,2030)) %>%
  fct_recode(`1990s` = "(1.99e+03,2e+03]",
             `2000s` = "(2e+03,2.01e+03]",
             `2010s` = "(2.01e+03,2.02e+03]",
             `2020s` = "(2.02e+03,2.03e+03]")

BATS_dsub$year <- cut(BATS_dsub$decy,breaks = round(min(BATS_dsub$decy)):ceiling(max(BATS_dsub$decy))) 

BATS_dsub_summary <- BATS_dsub %>%
  group_by(decade,nominal_depth) %>%
    summarise(across(
    .cols = c(Temp, DIC,pH,CO2,pCO2,HCO3,CO3,ALK),  # Add your columns here
    .fns = list(mean = ~mean(.x, na.rm = TRUE),
                sd = ~sd(.x, na.rm = TRUE)),
    .names = "{.col}_{.fn}"
  ))

lm(Temp ~ decy, BATS_dsub %>% subset(nominal_depth == 0 & decy < 2019))

BATS_dsub_summary_y <- BATS_dsub %>%
  group_by(year,nominal_depth) %>%
    summarise(across(
    .cols = c(Temp, DIC,pH,CO2,pCO2,HCO3,CO3,ALK),  # Add your columns here
    .fns = list(mean = ~mean(.x, na.rm = TRUE),
                sd = ~sd(.x, na.rm = TRUE)),
    .names = "{.col}_{.fn}"
  ))

BATS_dsub %>% subset(nominal_depth %in% c(0,40,100)) %>% ggplot() + geom_point(aes(month,Temp,color = factor(nominal_depth)))

BATS_dsub %>% subset(nominal_depth %in% c(0,40,100)) %>% ggplot() + geom_point(aes(month,HCO3,color = factor(nominal_depth)))

# clean up
BATS <- list(dataframe = BATS_dsub, means = BATS_dsub_summary, year_means = BATS_dsub_summary_y)
remove(BATS_carb,BATS_dsub,BATS_dsub_summary,sel,t,a)
gc()

```

CMIP-6 Abs and delta calculation
```{r}
# Pre-split the SpatVector by biome, ignore ice regions 
biome_list <- split(biomes_vec, biomes_vec$MeanBiomes)
biome_list <- biome_list[which(!names(biome_list) %in% c("NA_ICE","NP_ICE","SO_ICE"))]

index_model <- names(cmip6)
index_sen   <- names(cmip6[[1]])
index_vars  <- names(cmip6[[1]][[1]])
index_zone  <- names(biome_list)

fast_spatraster_to_long_dt <- function(x) {
  dt <- as.data.table(x, xy = TRUE)
  id_vars      <- c("x", "y")
  measure_vars <- setdiff(names(dt), id_vars)
  dt <- melt(dt, id.vars = id_vars, measure.vars = measure_vars,
             variable.name = "key", value.name = "value")
  dt[, c("variable", "month") := tstrsplit(key, "_(?=[^_]+$)", perl = TRUE)]
  dt[, key := NULL]
  return(dt)
}

add_period <- function(dt) {
  dt[month %in% as.character(1:12),  period := "1850s"]
  dt[month %in% as.character(13:24), period := "2000s"]
  dt[month %in% as.character(25:36), period := "2090s"]
  dt[, month := as.character(((as.integer(month) - 1) %% 12) + 1)]
  dt
}

key_cols <- c("x", "y", "month", "model", "variable")

# Initialise output containers
cmip6_df_means <- list(
  model       = list(),
  year        = list(),
  spatial     = list(),
  total       = list(),
  total_bloom = list(),
  seasonal    = list(),
  keep_spat   = list()
)

cmip6_df_anomaly_means <- list(
  total    = list(),
  seasonal = list()
)

for (sen in index_sen) {
  for (vars in index_vars) {
    for (zone in index_zone) {

      # --- 1. Build combined data.table across models ---
      model_chunks <- vector("list", length(index_model))
      names(model_chunks) <- index_model

      for (model in index_model) {
        cat(sen, model, vars, zone, "\n")

        r  <- mask(cmip6[[model]][[sen]][[vars]], biome_list[[zone]])
        dt <- fast_spatraster_to_long_dt(r)
        rm(r); gc()

        dt <- add_period(dt)
        dt[, model := model]
        dt <- dt[w_df, on = c("x", "y"), nomatch = 0]
        dt <- dt[!is.na(variable)]

        model_chunks[[model]] <- dt
        rm(dt)
      }

      combined <- data.table::rbindlist(model_chunks, use.names = TRUE, fill = TRUE)
      rm(model_chunks); gc()

      # --- 2. Regular aggregations ---

      cmip6_df_means$model[[sen]][[vars]][[zone]] <- combined[, .(
        model_mean = mean(value, na.rm = TRUE),
        model_sd   = sd(value, na.rm = TRUE),
        model_min  = min(value, na.rm = TRUE),
        model_max  = max(value, na.rm = TRUE),
        model_se   = sd(value, na.rm = TRUE) / sqrt(.N)
      ), by = .(x, y, month, period)]

      cmip6_df_means$year[[sen]][[vars]][[zone]] <- combined[, .(
        year_mean = mean(value, na.rm = TRUE),
        year_sd   = sd(value, na.rm = TRUE),
        year_min  = min(value, na.rm = TRUE),
        year_max  = max(value, na.rm = TRUE),
        year_se   = sd(value, na.rm = TRUE) / sqrt(.N)
      ), by = .(x, y, model, period)]

      cmip6_df_means$spatial[[sen]][[vars]][[zone]] <- combined[, .(
        spat_mean       = mean(value, na.rm = TRUE),
        spat_areaW_mean = weighted.mean(value, area_km2, na.rm = TRUE),
        spat_sd         = sd(value, na.rm = TRUE),
        spat_min        = min(value, na.rm = TRUE),
        spat_max        = max(value, na.rm = TRUE),
        spat_se         = sd(value, na.rm = TRUE) / sqrt(.N)
      ), by = .(model, month, period)]

      cmip6_df_means$total[[sen]][[vars]][[zone]] <- combined[, .(
        mean       = mean(value, na.rm = TRUE),
        areaW_mean = weighted.mean(value, area_km2, na.rm = TRUE),
        sd         = sd(value, na.rm = TRUE),
        min        = min(value, na.rm = TRUE),
        max        = max(value, na.rm = TRUE),
        se         = sd(value, na.rm = TRUE) / sqrt(.N)
      ), by = .(period)]

      if (zone %in% bloom_months$biome) {
        bloom <- unlist(bloom_months[bloom_months$biome == zone, "month"][[1]])
        cmip6_df_means$total_bloom[[sen]][[vars]][[zone]] <- combined[month %in% bloom, .(
          mean       = mean(value, na.rm = TRUE),
          areaW_mean = weighted.mean(value, area_km2, na.rm = TRUE),
          sd         = sd(value, na.rm = TRUE),
          min        = min(value, na.rm = TRUE),
          max        = max(value, na.rm = TRUE),
          se         = sd(value, na.rm = TRUE) / sqrt(.N)
        ), by = .(period)]
      }

      cmip6_df_means$seasonal[[sen]][[vars]][[zone]] <- combined[, .(
        seas_mean       = mean(value, na.rm = TRUE),
        seas_areaW_mean = weighted.mean(value, area_km2, na.rm = TRUE),
        seas_sd         = sd(value, na.rm = TRUE),
        seas_min        = min(value, na.rm = TRUE),
        seas_max        = max(value, na.rm = TRUE),
        seas_se         = sd(value, na.rm = TRUE) / sqrt(.N)
      ), by = .(month, period)]

      cmip6_df_means$keep_spat[[sen]][[vars]][[zone]] <- combined[, .(
        keep_spat_mean = mean(value, na.rm = TRUE),
        keep_spat_sd   = sd(value, na.rm = TRUE),
        keep_spat_min  = min(value, na.rm = TRUE),
        keep_spat_max  = max(value, na.rm = TRUE),
        keep_spat_se   = sd(value, na.rm = TRUE) / sqrt(.N)
      ), by = .(x, y, period, area_km2)]

      # --- 3. Anomalies ---
      # Extract 2000s baseline once and reuse for both anomaly comparisons
      dt_2000s <- combined[period == "2000s", .SD, .SDcols = c(key_cols, "value")]
      setnames(dt_2000s, "value", "value_2000s")

      # Merge baseline into 1850s and 2090s, compute anomalies, bind
      a_1850 <- merge(combined[period == "1850s"], dt_2000s, by = key_cols)
      a_1850[, `:=`(value = value - value_2000s, period = "1850s-2000s")][, value_2000s := NULL]

      a_2090 <- merge(combined[period == "2090s"], dt_2000s, by = key_cols)
      a_2090[, `:=`(value = value - value_2000s, period = "2090s-2000s")][, value_2000s := NULL]

      rm(dt_2000s)

      anomaly <- rbind(a_1850, a_2090)
      rm(a_1850, a_2090)

      # Anomaly aggregations
      cmip6_df_anomaly_means$total[[sen]][[vars]][[zone]] <- anomaly[, .(
        mean       = mean(value, na.rm = TRUE),
        areaW_mean = weighted.mean(value, area_km2, na.rm = TRUE),
        sd         = sd(value, na.rm = TRUE),
        min        = min(value, na.rm = TRUE),
        max        = max(value, na.rm = TRUE),
        se         = sd(value, na.rm = TRUE) / sqrt(.N)
      ), by = .(period)]

      cmip6_df_anomaly_means$seasonal[[sen]][[vars]][[zone]] <- anomaly[, .(
        seas_mean       = mean(value, na.rm = TRUE),
        seas_areaW_mean = weighted.mean(value, area_km2, na.rm = TRUE),
        seas_sd         = sd(value, na.rm = TRUE),
        seas_min        = min(value, na.rm = TRUE),
        seas_max        = max(value, na.rm = TRUE),
        seas_se         = sd(value, na.rm = TRUE) / sqrt(.N)
      ), by = .(month, period)]

      rm(anomaly, combined); gc()
    }
  }
}

# Clean up source data
rm(cmip6, w_df, weights); gc()

# Drop NULL entries from total_bloom
cmip6_df_means$total_bloom <- cmip6_df_means$total_bloom |>
  modify_depth(3, ~ if (is.atomic(.x) && length(.x) == 1 && is.na(.x)) NULL else .x) |>
  modify_depth(2, discard, is.null) |>
  modify_depth(1, discard, is.null)

flatten_for_plot <- function(x){
    # Flatten to a single list
    x <- list_flatten(x) %>% list_flatten()
    
    # Names are like "ssp126.dic.zone1" — split them
    x <- as.data.frame(data.table::rbindlist(x, idcol = "id"))
    
    matches <- str_match(x$id, "^([^_]+)_([^_]+)_(.+)$")
    x$scenario <- matches[, 2]
    x$var <- matches[, 3]
    x$biome <- matches[, 4]
    
    matches <- str_match(x$biome, "^([^_]+)_([^_]+)")
    x$loc <- matches[, 2]
    x$type <- matches[, 3]
    
    remove(matches)
    return(x)
}

```

Lab Data - DIC and pH
```{r}
library(readxl)
library(ldeomeb)

dic <- readxl::read_xlsx("C:/Users/henryholm/OneDrive - Woods Hole Oceanographic Institution/Desktop/ORCC/Copy of HurleyS 250311 ORCC_250303 P240998.xlsx",sheet = 2)
dic2 <- readxl::read_xlsx("C:/Users/henryholm/OneDrive - Woods Hole Oceanographic Institution/Desktop/ORCC/HurleyS 250730 DIP ORCC P262351.xlsx",sheet = 2)

# load pH plate data
loc <- "C:/Users/henryholm/OneDrive - Woods Hole Oceanographic Institution/Desktop/ORCC/"
loc <- paste0(loc, c("ORCC_pH_Complete_Book.xlsx", "ORCC_pH_Complete_Book_preJan13th.xlsx"))
# import data from all sheets and combine
all_sheets <- lapply(loc, function(loc) {
  # Get all sheet names
  sheet_names <- excel_sheets(loc)

  # Read each sheet into a named list of data.frames
  all_sheets <- lapply(sheet_names, read_excel, path = loc)
  names(all_sheets) <- sheet_names
  return(all_sheets)
}) %>%
  list_flatten()

# drop any empty sheets
all_sheets <- all_sheets[!sapply(all_sheets, is_empty)]

# drop sheets without name data one without names
all_sheets <- all_sheets[which(!names(all_sheets) %in% c("Jan_1_22_2025","Apr_15_25"))]

ph_spec <- map(all_sheets, function(x){x %>% data.frame() %>% ldeomeb::tidyplate()},.progress = T) 

# drop this bad replicate
dic2 <- dic2[which(!dic2$`Sample ID` == "Tpsu_O417_A"),]

dic <- bind_rows(dic,dic2)

loc <- "https://docs.google.com/spreadsheets/d/16Oq4psIy7DFOsbAWLwy2RkPCYLFqkRjJey-ki3iN2SQ/edit"

# import
curves <- googlesheets4::read_sheet(loc,sheet = "Curves")
cultures <- googlesheets4::read_sheet(loc,sheet = "Cultures")
samples <- googlesheets4::read_sheet(loc,sheet = "Samples")

ph_spec <- ph_spec %>%
  map(~ .x$plates %>%
        subset(!is.na(value) | !is.na(names)) %>%
        calc_plate(tidyplate = .,pairs = .x$pairs),
      .progress = T) %>%
  bind_rows()

# fix names
ph_spec$`Culture_ID#` <- ph_spec$names %>%
  str_replace(pattern = "tpsu",replacement = "T_pseudo") %>%
  str_replace(pattern = "Tpsu",replacement = "T_pseudo") %>%
  str_replace(pattern = "7002",replacement = "Syn_7002") %>%
  str_replace(pattern = "Ehux",replacement = "E_huxleyi") %>%
  str_replace(pattern = "8102",replacement = "Syn_8102") %>%
  str_replace(pattern = "Syn_Syn",replacement = "Syn")

# create same date columns for matching
curves$date <- date(curves$Meaurement_DateTime)
ph_spec$date <- date(ph_spec$`Average Start Time`)

#rename and match
ph_spec <- ph_spec %>% rename(pH_total_spec = pH)
curves <- left_join(curves,ph_spec,by = join_by(`Culture_ID#`,date))

dic$`Culture_ID#` <- dic$`Sample ID Offical`

joined <- left_join(dic,curves)

joined <- samples %>%
          subset(!duplicated(`Sample Name`)) %>% 
  select(`Sample Name`,`Harvest Type`) %>%
  left_join(joined,.,by = join_by(`Sample ID Offical` == `Sample Name`))

for (i in seq_along(joined$pH_volt)) {
  if (is.na(joined[i, "pH_temp"])) {
    joined[i, "pH_total_volt"] <- NA
  } else {
    joined[i, "pH_total_volt"] <- seacarb::pH(
      Ex = joined[i, "pH_volt"],
      Etris = joined[i, "pH_std_volt"],
      T = joined[i, "pH_temp"]
    )
  }
}

lm(joined$pH_total_volt ~ joined$pH_NIST)

ggplot(joined,aes(pH_total_volt,pH_NIST)) +
  geom_point() +
  geom_abline() +
  geom_abline(slope = 1,intercept = 0.2,linetype = "dashed") +
  theme_minimal()

joined %>% subset(pH_total_volt > 6) %>% # drop one outlier
ggplot(aes(pH_total_volt)) +
  geom_density(aes(fill = Condition)) 

joined %>% subset(pH_total_volt > 6) %>% 
  group_by(Condition) %>%
  summarise(mean = mean(pH_total_volt))

joined %>% select(`Culture_ID#`,Meaurement_DateTime,pH_total_volt,pH_total_spec) %>%
  subset(complete.cases(.)) %>%
  ggplot(aes(pH_total_spec,pH_total_volt)) +
  geom_abline(aes(slope = 1, intercept = 0, linetype = "1:1 Line",color = "1:1 Line"), size =0.75) +
  stat_smooth(method = "lm",alpha = 0.1) +
  ggpubr::stat_regline_equation() +
  theme_minimal() +
  geom_point(shape = 1) +
  scale_linetype_manual(values = c("solid",'dotted',"solid")) +
  scale_color_manual(values = c("red","black","black")) +
  labs(x = "pH Total Scale (Voltage Based)",
       y = "pH Total Scale (m-cresol)",
       linetype = "", color = ""
       )

joined <- joined %>% subset(complete.cases(pH_total_volt,`DIC (mM)`,Temp.nominal))

curves %>% subset(pH_NIST > 5 & pH_NIST < 10) %>%
  group_by(Incubator.CO2.ppm.nominal) %>%
  summarise(`Total pH` = mean(pH_NIST-0.23))

### full carb calc
joined <- pyco2$sys(
  par1      = joined$pH_total_volt,
  par2      = joined$`DIC (mM)`/1.0265/1e3, #measured density w. 
  par1_type = 3L,
  par2_type = 2L,
  salinity  = 35,
  opt_pH_scale = 1,
  temperature = joined$Temp.nominal,
  pressure  = 0,
  opt_k_carbonic = 10,
  opt_k_bisulfate = 1,
  opt_total_borate =1,
  opt_k_fluoride = 1
) %>% 
  data.frame() %>% 
  rename_with(~ paste0("pyco2sys_", .)) %>%
  cbind(joined,.)
  
joined <- seacarb::carbfull(flag = 9,
                        var1 = joined$pH_total_volt,
                        var2 = joined$`DIC (mM)`/1.0265/1e3,
                        T = joined$Temp.nominal,
                        S = 35,
                        pHscale = "T",
                        k1k2 = "l", ks = "d", kf = "dg" # match PyCO2sys settings
                        ) %>% 
  rename_with(~ paste0("seacarb_", .)) %>%
  cbind(joined,.)

plot((joined$seacarb_CO2-joined$pyco2sys_CO2)*1e6)
plot((joined$seacarb_HCO3-joined$pyco2sys_HCO3)*1e6)

df <- stringr::str_split(joined$`Culture_ID#`,pattern = "_",simplify = TRUE)
  
joined$sp <- paste(df[,1],df[,2],sep = "_")

# just looking at Ehux and Tpsu for this paper
joined <- joined %>% subset(Name %in% c("Ehux", "Tpsu"))

joined %>% 
  group_by(Incubator.CO2.ppm.nominal) %>%
  summarise(`mean pCO2 (calculated)` = mean(seacarb_pCO2),
            `sd pCO2 (calculated)` = sd(seacarb_pCO2))

r <- joined %>%
    subset(Blank %in% c("Sample_Final") & !sp %in% c("NA_","7002")) %>%
    group_by(`Sample ID Offical`,Blank,Incubator.CO2.ppm.nominal,Name,`Harvest Type`,Medium) %>%
    summarise(
    across(
      where(is.numeric),
      .fns = list(
        mean = ~mean(unique(.x), na.rm = TRUE),
        sd   = ~sd(unique(.x), na.rm = TRUE)
      )
    )
  )
r$Name <- unlist(r$Name)

s <- joined %>%
    subset(Blank %in% c("Sample_Final") & !sp %in% c("NA_","7002")) %>%
    group_by(Incubator.CO2.ppm.nominal,Name) %>%
    summarise(
    across(
      where(is.numeric),
      .fns = list(
        mean = ~mean(unique(.x), na.rm = TRUE),
        sd   = ~sd(unique(.x), na.rm = TRUE)
      )
    )
  )

s$Name <- unlist(s$Name)


b <- joined %>%
  group_by(`Sample ID Offical`) %>%
  slice_min(order_by = `Growth Time Nominal (hours)`,n = 1,with_ties = TRUE) %>%
  subset(Blank %in% c("Blank_Initial") & sp != "NA_") %>% group_by(Name,Condition) %>%
    summarise(
    across(
      where(is.numeric),
      .fns = list(
        mean = ~mean(unique(.x), na.rm = TRUE),
        sd   = ~sd(unique(.x), na.rm = TRUE)
      )
    )
  )

b$Name <- unlist(b$Name)

t <- joined %>%
  subset(sp != "NA_") %>%
  group_by(Name,Incubator.CO2.ppm.nominal) %>%
  summarise(
    across(
      where(is.numeric),
      .fns = list(
        mean = ~mean(unique(.x), na.rm = TRUE),
        sd   = ~sd(unique(.x), na.rm = TRUE)
      )
    )
  )
t$Name <- unlist(t$Name)

a <- joined %>%
  subset(sp != "NA_" & Name != "7002") %>%
   group_by(Incubator.CO2.ppm.nominal) %>%
    summarise(
    across(
      where(is.numeric),
      .fns = list(
        mean = ~mean(unique(.x), na.rm = TRUE),
        sd   = ~sd(unique(.x), na.rm = TRUE)
      )
    )
  ) %>%
  mutate(Name = "All Sp")

ab <- joined %>%
  group_by(Incubator.CO2.ppm.nominal) %>%
  subset(Blank %in% c("Blank_Initial") & !sp %in% c("NA_","7002")) %>%
  #group_by(Incubator.CO2.ppm.nominal,Name) %>%
    summarise(
    across(
      where(is.numeric),
      .fns = list(
        mean = ~mean(unique(.x), na.rm = TRUE),
        sd   = ~sd(unique(.x), na.rm = TRUE)
      )
    )
  )%>%
  mutate(Name = "All Sp")
ab$Name <- unlist(ab$Name)

as <- joined %>%
  group_by(`Sample ID Offical`) |>
  subset(Blank %in% c("Sample_Final")) %>%
  slice_max(`Growth Time Nominal (hours)`, n = 1) %>%
  group_by(Incubator.CO2.ppm.nominal) %>%
    summarise(
    across(
      where(is.numeric),
      .fns = list(
        mean = ~mean(unique(.x), na.rm = TRUE),
        sd   = ~sd(unique(.x), na.rm = TRUE)
      )
    )
  )%>%
  mutate(Name = "All Sp")

as$Name <- unlist(as$Name)

joined %>%
  group_by(`Sample ID Offical`,Blank) %>%
    summarise(
    across(
      where(is.numeric),
      .fns = list(
        mean = ~mean(unique(.x), na.rm = TRUE),
        sd   = ~sd(unique(.x), na.rm = TRUE)
      )
    )
  )
  

lab <- bind_rows(list(reps = r,
                      samples = s, blanks = b, total = t, all = a, all_blank = ab, all_sample = as),.id = 'samp_group') 

current <- joined %>%
  subset(Name != "7002") 

lab[which(lab$Condition == "Ambient"),"temp"] <- 22
lab[which(lab$Condition == "1135ppm"),"temp"] <- 25.3

remove(df,s,t,b,ab,as,r)

```

### Plotting Section ###

Figure 1 - Delta Scaling
```{r}
flatten_for_plot <- function(x){
    # Flatten to a single list
    x <- list_flatten(x) %>% list_flatten()
    
    # Names are like "ssp126.dic.zone1" — split them
    x <- as.data.frame(data.table::rbindlist(x, idcol = "id"))
    
    matches <- str_match(x$id, "^([^_]+)_([^_]+)_(.+)$")
    x$scenario <- matches[, 2]
    x$var <- matches[, 3]
    x$biome <- matches[, 4]
    
    matches <- str_match(x$biome, "^([^_]+)_([^_]+)")
    x$loc <- matches[, 2]
    x$type <- matches[, 3]
    
    remove(matches)
    return(x)
}

# Totals Plot

delta <- flatten_for_plot(cmip6_df_anomaly_means$total)
abs <- flatten_for_plot(cmip6_df_means$total)

delta[which(delta$period == "1850s-2000s"),"period"] <- "1850s"
delta[which(delta$period == "2090s-2000s"),"period"] <- "2090s"

delta <- full_join(delta,abs,
                   by = join_by(id,period,scenario,var,biome,loc,type),
                   suffix = c("_delta","_abs")
                   )

delta <- delta %>%
  pivot_wider(
    id_cols = c(scenario, biome, loc, type, period),
    names_from = var, values_from = c(areaW_mean_delta, sd_delta,areaW_mean_abs,sd_abs,min_delta,max_delta)
  )

delta$type <- factor(delta$type,levels = c("EQU","STPS","STSS","SPSS","ICE"))

delta$p_s <-  paste(delta$scenario, delta$period,  sep = " - ")

delta[which(delta$p_s %in% c("ssp245 - 1850s","ssp585 - 1850s")),"p_s"] <- "Historical , 1850s"
delta[which(delta$p_s %in% c("ssp245 - 1850s","ssp585 - 1850s")),"p_s"] <- "Historical , 1850s"
delta[which(delta$p_s %in% c("ssp585 - 2090s")),"p_s"] <- "SSP5-8.5 , 2090s"
delta[which(delta$p_s %in% c("ssp245 - 2090s")),"p_s"] <- "SSP2-4.5 , 2090s"

plot_delta <- function(data, var_x_suffix, var_y_suffix, y_lab = NA, x_lab = NA) {
  
  # Check for missing or NA character labels
  if (missing(y_lab) || (is.character(y_lab) && length(y_lab) == 1 && is.na(y_lab))) {
    y_lab <- var_y_suffix
  }
  if (missing(x_lab) || (is.character(x_lab) && length(x_lab) == 1 && is.na(x_lab))) {
    x_lab <- var_x_suffix
  }
  
  x_col <- paste0("areaW_mean_delta_", var_x_suffix)
  y_col <- paste0("areaW_mean_delta_", var_y_suffix)
  sd_x_col <- paste0("sd_delta_", var_x_suffix)
  sd_y_col <- paste0("sd_delta_", var_y_suffix)
  
  x_sym <- rlang::sym(x_col)
  y_sym <- rlang::sym(y_col)
  sd_x_sym <- rlang::sym(sd_x_col)
  sd_y_sym <- rlang::sym(sd_y_col)
  
  ggplot(
    data %>% subset(type != "ICE" & period != "2000s"),
    aes(x = !!x_sym, y = !!y_sym, color = p_s,shape = type)
  ) +
    geom_hline(yintercept = 0, color = "black", linewidth = 0.1, linetype = "dashed") +
    geom_vline(xintercept = 0, color = "black", linewidth = 0.1, linetype = "dashed") +
    geom_point(size = 2) +
    geom_errorbar(aes(
      ymin = !!y_sym - !!sd_y_sym,
      ymax = !!y_sym + !!sd_y_sym
    ),linewidth = 0.1) +
    geom_errorbarh(aes(
      xmin = !!x_sym - !!sd_x_sym,
      xmax = !!x_sym + !!sd_x_sym
    ),linewidth = 0.1) +
    labs(
      y = if (is.expression(y_lab)) y_lab else paste0(y_lab),
      x = if (is.expression(x_lab)) x_lab else paste0(x_lab),
      color = "Period and Scenario", shape = "Biome Type"
    ) +
    scale_color_manual(values = c("grey50", "orange", "maroon")) +
    scale_shape_manual(values = c(15:17,4))
}

dhco3 <- expression(paste("Δ [HCO"[3]^"-","]", " (", "umol ", kg^{-1}, ")"))
dco2 <- expression("Δ [CO"[2]*"] (umol kg"^{-1}*")")
dco3 <- expression(paste("Δ [CO"[3]^"-2","]", " (", "umol ", kg^{-1}, ")"))
dtemp <- "Δ Temperature (°C)"
dpco2 <- expression("Δ pCO"[2]*"")

legend <- ggpubr::get_legend(plot_delta(delta,"tos","pCO2",x_lab = dtemp,y_lab = dpco2) + theme(legend.position = "top"))


fig1_bottom <- (
  plot_delta(delta,"tos","pCO2", x_lab = dtemp, y_lab = dpco2) +
    plot_delta(delta,"aqueous","pCO2", x_lab = dco2, y_lab = dpco2) + 
    plot_delta(delta,"bicarbonate","pCO2", x_lab = dhco3, y_lab = dpco2) + 
    plot_delta(delta,"carbonate","pCO2", x_lab = dco3, y_lab = dpco2) +
    plot_delta(delta,"pH","pCO2", x_lab = "Δ pH (Total Scale)", y_lab = dpco2) +
    plot_delta(delta,"dic","pCO2", x_lab = expression("Δ DIC (umol kg"^{-1}*")"), y_lab = dpco2) 
) +
  plot_layout(guides = "collect", nrow = 1, axes = "collect") &
  #plot_annotation(tag_levels = list(c("G", "H", "I","J", "K", "L"))) 
  theme(
    legend.position = "none",
    plot.tag = element_text(
      face = "bold",
      #family = "Arial",
      size  = 12,
      hjust = -0.1,
      vjust = 1.1
    )
  )

```

Figure 1 - Absolute Conditions
```{r}
abs <- flatten_for_plot(cmip6_df_means$total)

abs <- abs %>%
  pivot_wider(
    id_cols = c(scenario, biome, loc, type, period),
    names_from = var, values_from = c(areaW_mean, sd, min, max)
  ) 

abs <- delta

abs$type <- factor(abs$type,levels = c("EQU","STPS","STSS","SPSS","ICE"))

abs$p_s <-  paste(abs$scenario, abs$period,  sep = " - ")

abs[which(abs$p_s %in% c("ssp245 - 1850s","ssp585 - 1850s")),"p_s"] <- "Historical , 1850s"
abs[which(abs$p_s %in% c("ssp245 - 1850s","ssp585 - 1850s")),"p_s"] <- "Historical , 1850s"
abs[which(abs$p_s %in% c("ssp585 - 2090s")),"p_s"] <- "SSP5-8.5 , 2090s"
abs[which(abs$p_s %in% c("ssp245 - 2090s")),"p_s"] <- "SSP2-4.5 , 2090s"
abs[which(abs$p_s %in% c("ssp585 - 2000s")),"p_s"] <- "Modern , 2000s"
abs[which(abs$p_s %in% c("ssp245 - 2000s")),"p_s"] <- "Modern , 2000s"

abs$p_s <- factor(abs$p_s,levels = c("Historical , 1850s",
                                     "Modern , 2000s",
                                     "SSP2-4.5 , 2090s",
                                     "SSP5-8.5 , 2090s"))


plot_abs_data <- function(data, 
                          xvar, yvar,
                          xlab = NULL, ylab = NULL,sdevs = 1) {
  
  # Capture variable names
  x_name <- rlang::as_name(rlang::enquo(xvar))
  y_name <- rlang::as_name(rlang::enquo(yvar))
  
  # Auto-find matching SD columns (assumes "sd_" prefix)
  sd_x_name <- paste0("sd_", sub("^areaW_mean_", "", x_name))
  sd_y_name <- paste0("sd_", sub("^areaW_mean_", "", y_name))
  
  # Check if they exist, else set NULL
  sd_x <- if (sd_x_name %in% names(data)) rlang::sym(sd_x_name) else NULL
  sd_y <- if (sd_y_name %in% names(data)) rlang::sym(sd_y_name) else NULL
  
  ggplot_data <- data %>%
    subset(type != "ICE") %>%
    group_by(scenario, biome, loc, type, period) %>%
    ungroup()
  
  p <- ggplot(ggplot_data, aes(x = {{ xvar }}, y = {{ yvar }},
                               color = p_s, shape = type)) +
    #geom_hline(yintercept = 0, color = "black", linewidth = 0.1, linetype = "dashed") +
    #geom_vline(xintercept = 0, color = "black", linewidth = 0.1, linetype = "dashed") +
    geom_point()
  
  # Add error bars if found
  if (!is.null(sd_y)) {
    p <- p + geom_errorbar(aes(ymin = {{ yvar }} - !!sd_y*sdevs,
                               ymax = {{ yvar }} + !!sd_y*sdevs),
                           linewidth = 0.1)
  }
  if (!is.null(sd_x)) {
    p <- p + geom_errorbar(aes(xmin = {{ xvar }} - !!sd_x*sdevs,
                               xmax = {{ xvar }} + !!sd_x*sdevs),
                           linewidth = 0.1)
  }
  
  p +
    labs(
      x = ifelse(is.null(xlab), x_name, xlab),
      y = ifelse(is.null(ylab), y_name, ylab),
      color = "Period and Scenario", 
      shape = "Biome Type"
    ) +
    scale_color_manual(values = c("grey50","#3295a8","orange","maroon")) +
    scale_shape_manual(values = c(15:17,4))
}

hco3 <- expression(paste("[HCO"[3]^"-","]", " (", "umol ", kg^{-1}, ")"))
co2 <- expression("[CO"[2]*"] (umol kg"^{-1}*")")
co3 <- expression(paste("[CO"[3]^"-2","]", " (", "umol ", kg^{-1}, ")"))
temp <- "Temperature (°C)"
pco2 <- expression("pCO"[2]*"")

fig1_top <- (
  plot_abs_data(abs,
    xvar = areaW_mean_abs_tos, yvar = areaW_mean_abs_pCO2,
    xlab = temp, ylab = pco2
  ) + theme(legend.position = "none") +
  plot_abs_data(abs,
    xvar = areaW_mean_abs_aqueous, yvar = areaW_mean_abs_pCO2,
    xlab = co2, ylab = pco2
  ) + theme(legend.position = "none") +
  plot_abs_data(abs,
    xvar = areaW_mean_abs_bicarbonate, yvar = areaW_mean_abs_pCO2,
    xlab = hco3, ylab = pco2
  ) + theme(legend.position = "none") +
  plot_abs_data(abs,
    xvar = areaW_mean_abs_carbonate, yvar = areaW_mean_abs_pCO2,
    xlab = co3, ylab = pco2
  ) + theme(legend.position = "none") +
  plot_abs_data(abs,
    xvar = areaW_mean_abs_pH, yvar = areaW_mean_abs_pCO2,
    xlab = "pH (Total Scale)", ylab = pco2
  ) + theme(legend.position = "none") +
  plot_abs_data(abs,
    xvar = areaW_mean_abs_dic, yvar = areaW_mean_abs_pCO2,
    xlab = expression("DIC (umol kg"^{-1}*")"), ylab = pco2
  ) + theme(legend.position = "none") + scale_x_continuous(n.breaks = 4) 
) +
  plot_layout(guides = "collect", nrow = 1, axes = "collect") &
  #plot_annotation(tag_levels = list(c("A","B","C","D","E","F"))) 
  theme(
    plot.tag = element_text(
      face = "bold",
     # family = "Arial", 
      size  = 12,
      hjust = -0.1,
      vjust = 1.1
    )
  )

fig1_ab_supp <- (
  plot_abs_data(abs,
    xvar = areaW_mean_pH, yvar = areaW_mean_pCO2,
    xlab = "pH (Total Scale)", ylab = pco2
  ) + theme(legend.position = "none") +
  plot_abs_data(abs,
    xvar = areaW_mean_dic, yvar = areaW_mean_pCO2,
    xlab = expression("DIC (umol kg"^{-1}*")"), ylab = pco2
  ) + theme(legend.position = "none") 
) +
  plot_layout(guides = "collect", nrow = 1, axes = "collect") +
  #plot_annotation(tag_levels = list(c("A","B"))) 
  theme(
    plot.tag = element_text(
      face = "bold",
      #family = "Arial",  
      size  = 12,
      hjust = -0.1,
      vjust = 1.1
    )
  )

(fig1_top / fig1_bottom) + plot_annotation(tag_levels = "A")

```

Figure 1 - Complete Plot
```{r}
fig1_bcde / fig1_fghi + plot_annotation(tag_levels = list(c("b","c","d","e","f","g","h","i"))) 

fig1_ab_supp / fig1_cd_supp + plot_annotation(tag_levels = list(c("A","B","C","D"))) 

```

Figure 3 - Delta Plots
```{r}

plot_delta(delta,"bicarbonate","aqueous", x_lab = dhco3, y_lab = dco2)

plot_abs_data(abs,
    xvar = areaW_mean_aqueous, yvar = areaW_mean_bicarbonate,
    ylab = hco3, xlab = co2)
  

```

GLODAP Check Supp Fig
```{r}
compare <- abs %>% subset(period == "2000s" & type != "ICE") %>%
  left_join(rename_all(glodap$means,.funs = ~ paste0(.x,"_glodap")),join_by(biome == "loc_glodap")) 

compare %>%
  ggplot(aes(areaW_mean_tos,mean_temp_glodap)) +
  geom_point(aes(shape = type)) +
  geom_abline(linetype = "dashed") + 
  geom_errorbar(aes(ymax = mean_temp_glodap + errors::errors(mean_temp_glodap),
                    ymin = mean_temp_glodap - errors::errors(mean_temp_glodap))) +
    geom_errorbar(aes(xmax = areaW_mean_tos + sd_tos,
                    xmin = areaW_mean_tos - sd_tos)) +
    scale_shape_manual(values = c(15:17,4)) +
  labs(y = "GLODAP - Annual Mean Temperature (°C)",
       x = "CMIP6 - Annual Mean Temperature (°C)") +

compare %>%
  ggplot(aes(areaW_mean_dic,mean_TCO2_glodap)) +
  geom_point(aes(shape = type)) +
  geom_abline(linetype = "dashed") + 
  geom_errorbar(aes(ymax = mean_TCO2_glodap + errors::errors(mean_TCO2_glodap),
                    ymin = mean_TCO2_glodap - errors::errors(mean_TCO2_glodap))) +
    geom_errorbar(aes(xmax = areaW_mean_dic + sd_dic,
                    xmin = areaW_mean_dic - sd_dic)) +
    scale_shape_manual(values = c(15:17,4)) +
  labs(y = expression(paste("GLODAP - Annual Mean DIC"," (", "umol ", kg^{-1}, ")")),
       x = expression(paste("CMIP6 - Annual Mean DIC"," (", "umol ", kg^{-1}, ")"))) +
  
compare %>%
  ggplot(aes(areaW_mean_pH,mean_pH_glodap)) +
  geom_point(aes(shape = type)) +
  geom_abline(linetype = "dashed") + 
  geom_errorbar(aes(ymax = mean_pH_glodap + errors::errors(mean_pH_glodap),
                    ymin = mean_pH_glodap - errors::errors(mean_pH_glodap))) +
    geom_errorbar(aes(xmax = areaW_mean_pH + sd_pH,
                    xmin = areaW_mean_pH - sd_pH)) +
    scale_shape_manual(values = c(15:17,4)) +
  labs(y = expression(paste("GLODAP - Annual Mean DIC"," (", "umol ", kg^{-1}, ")")),
       x = expression(paste("CMIP6 - Annual Mean DIC"," (", "umol ", kg^{-1}, ")"))) 


ggplot() + geom_density(aes(compare$mean_pH_glodap-compare$areaW_mean_pH)) 

```

SODA Check Supp Fig
```{r}
SODA_total


SODA_total <- SODA_total %>%
  separate(biome,into = c("loc","type"),"_",remove = F,extra = "merge") %>%
  pivot_wider(
    id_cols = c(biome, loc, type),
    names_from = variable , values_from = c(areaW_mean, sd, min, max)
  ) 

compare <- abs %>% subset(period == "2000s" & type != "ICE") %>%
  left_join(rename_all(SODA_total,.funs = ~ paste0(.x,"_soda")),join_by(biome == biome_soda)) 

make_comparison_plot <- function(df, x, sd_x, y, sd_y, x_lab, y_lab) {
  ggplot(df, aes(.data[[x]], .data[[y]])) +
    geom_point(aes(shape = type)) +
    geom_abline(linetype = "dashed") +
    geom_errorbar(aes(
      xmin = .data[[x]] - .data[[sd_x]],
      xmax = .data[[x]] + .data[[sd_x]]
    )) +
    geom_errorbar(aes(
      ymin = .data[[y]] - .data[[sd_y]],
      ymax = .data[[y]] + .data[[sd_y]]
    )) +
    scale_shape_manual(values = c(15:17, 4)) +
    labs(x = x_lab, y = y_lab)
}

compare %>%
  ggplot(aes(areaW_mean_tos,areaW_mean_temperature_soda)) +
  geom_point(aes(shape = type)) +
  geom_abline(linetype = "dashed") + 
    geom_errorbar(aes(xmax = areaW_mean_tos + sd_tos,
                    xmin = areaW_mean_tos - sd_tos)) +
    geom_errorbar(aes(ymax = areaW_mean_temperature_soda + sd_temperature_soda,
                    ymin = areaW_mean_temperature_soda - sd_temperature_soda)) +
    scale_shape_manual(values = c(15:17,4)) +
  labs(y = "SODA - Temperature (°C)",
       x = "CMIP6 - Temperature (°C)") +
compare %>%
  ggplot(aes(areaW_mean_aqueous,areaW_mean_co2_soda)) +
  geom_point(aes(shape = type)) +
  geom_abline(linetype = "dashed") + 
    geom_errorbar(aes(xmax = areaW_mean_aqueous + sd_aqueous,
                    xmin = areaW_mean_aqueous - sd_aqueous)) +
    geom_errorbar(aes(ymax = areaW_mean_co2_soda + sd_co2_soda,
                    ymin = areaW_mean_co2_soda - sd_co2_soda)) +
    scale_shape_manual(values = c(15:17,4)) +
  labs(y = expression(paste("SODA - [CO"[2]*"] (umol kg"^{-1}*")")),
       x = expression(paste("CMIP6 - [CO"[2]*"] (umol kg"^{-1}*")"))) +
  plot_layout(guides = "collect") +
compare %>%
  ggplot(aes(areaW_mean_bicarbonate,areaW_mean_hco3_soda)) +
  geom_point(aes(shape = type)) +
  geom_abline(linetype = "dashed") + 
    geom_errorbar(aes(xmax = areaW_mean_bicarbonate + sd_bicarbonate,
                    xmin = areaW_mean_bicarbonate - sd_bicarbonate)) +
    geom_errorbar(aes(ymax = areaW_mean_hco3_soda + sd_hco3_soda,
                    ymin = areaW_mean_hco3_soda - sd_hco3_soda)) +
    scale_shape_manual(values = c(15:17,4)) +
  labs(y = expression(paste("SODA - [HCO"[3]^"-","]", " (", "umol ", kg^{-1}, ")")),
       x = expression(paste("CMIP6 - [HCO"[3]^"-","]", " (", "umol ", kg^{-1}, ")"))) +
  plot_layout(guides = "collect") +
  compare %>%
  ggplot(aes(areaW_mean_carbonate,areaW_mean_co3_soda)) +
  geom_point(aes(shape = type)) +
  geom_abline(linetype = "dashed") + 
    geom_errorbar(aes(xmax = areaW_mean_carbonate + sd_carbonate,
                    xmin = areaW_mean_carbonate - sd_carbonate)) +
    geom_errorbar(aes(ymax = areaW_mean_co3_soda + sd_co3_soda,
                    ymin = areaW_mean_co3_soda - sd_co3_soda)) +
    scale_shape_manual(values = c(15:17,4)) +
  labs(y = expression(paste("SODA - [CO"[3]^"2-","]", " (", "umol ", kg^{-1}, ")")),
       x = expression(paste("CMIP6 - [CO"[3]^"2-","]", " (", "umol ", kg^{-1}, ")"))) +
  plot_layout(guides = "collect") +
  compare %>%
ggplot(aes(areaW_mean_dic,areaW_mean_dic_soda)) +
  geom_point(aes(shape = type)) +
  geom_abline(linetype = "dashed") + 
    geom_errorbar(aes(xmax = areaW_mean_dic + sd_dic,
                    xmin = areaW_mean_dic - sd_dic)) +
    geom_errorbar(aes(ymax = areaW_mean_dic_soda + sd_dic_soda,
                    ymin = areaW_mean_dic_soda - sd_dic_soda)) +
    scale_shape_manual(values = c(15:17,4)) +
  labs(y = expression(paste("SODA - DIC", " (", "umol ", kg^{-1}, ")")),
       x = expression(paste("CMIP6 - DIC", " (", "umol ", kg^{-1}, ")"))) +
  plot_layout(guides = "collect") +
compare %>%
  ggplot(aes(areaW_mean_pH,areaW_mean_ph_total_soda)) +
  geom_point(aes(shape = type)) +
  geom_abline(linetype = "dashed") + 
    geom_errorbar(aes(xmax = areaW_mean_pH + sd_pH,
                    xmin = areaW_mean_pH - sd_pH)) +
    geom_errorbar(aes(ymax = areaW_mean_ph_total_soda + sd_ph_total_soda,
                    ymin = areaW_mean_ph_total_soda - sd_ph_total_soda)) +
    scale_shape_manual(values = c(15:17,4)) +
  labs(y = "SODA - pH (Total Scale)",
       x = "CMIP6 - pH (Total Scale)") + plot_layout(guides = "collect")  + plot_annotation(tag_levels = "A")

compare$areaW_mean_tos-compare$areaW_mean_temperature_soda
compare$areaW_mean_aqueous/compare$areaW_mean_co2_soda
compare$areaW_mean_bicarbonate/compare$areaW_mean_hco3_soda
compare$areaW_mean_carbonate/compare$areaW_mean_co3_soda
compare$areaW_mean_pH/compare$areaW_mean_ph_total_soda

```


Figure 3 - DIC Alignment with NA-STPS and lab replication
```{r}
# pH Panel

# convert pvalue to significance level
signif.num <- function(x) {
    symnum(x, corr = FALSE, na = FALSE, legend = FALSE,
           cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1), 
           symbols = c("<0.001", "<0.01", "<0.05", "<0.1", "NonSig"))
}

sp <- c("Ehux","Tpsu")
formal <- c(Ehux = "G. huxleyi", Tpsu = "T. pseudonana")

ph_cd <- list()

# dfs
for (i in 1:length(sp)) {
  ph_cd$dfs[[sp[i]]] <- pHcd_all %>%
    subset(Species == sp[i] &
      complete.cases(.))
}

# models
for (i in 1:length(sp)) {
ph_cd$exp_mods[[sp[i]]]<- ph_cd$dfs[[sp[i]]] %>%
  mutate(lnCellDensity = log10(CellDensity), lnpH_offset = log10(pH_offset)) %>%
  lm(lnpH_offset ~ lnCellDensity,data = .)
}

ph_cd_param <- map_dfr(sp, function(species) {
  mod <- ph_cd[["exp_mods"]][[species]]
  mod_summary <- summary(mod)
  
  tibble(
    Species = species,
    Intercept = coef(mod)[1],
    Slope = coef(mod)[2],
    `Adj. R^2` = mod_summary$adj.r.squared,
    `Slope p-value` = signif.num(mod_summary$coefficients[2, 4]),
    `Intercept p-value` = signif.num(mod_summary$coefficients[1, 4]),
  )
})

ph_cd_param %>% gt::gt() %>% gtExtras::gt_theme_pff()

ph <- seq(0,2,0.01)
names(ph_cd_param$Intercept) <- ph_cd_param$Species
ph <- cbind(ph,as.data.frame(map2(ph_cd_param$Intercept,ph_cd_param$Slope,~ 10^((log10(ph) - .x)/.y)))) %>%
  pivot_longer(ph_cd_param$Species,names_to = 'sp',values_to = "cd")
  
for (i in 1:length(sp)) {
  ph_offset <- 0.1
  cells_ml <- as.numeric(ph %>% subset(sp == sp[i] & ph == ph_offset) %>% select(cd))
  ph_cd[["plots"]][[sp[i]]] <- ph_cd$dfs[[sp[i]]] %>%
  ggplot() +
  geom_point(aes(CellDensity,pH_offset,
                 #color = container,shape = as.character(Incubator.CO2.ppm.nominal)
                 ),size = 2) +
  geom_line(aes(cd,ph),data = subset(ph,sp == sp[i]),size = 0.5) +
  annotate(geom = "segment",y = ph_offset,yend = ph_offset,x = 0,xend = cells_ml,linetype = "dashed") +
  annotate(geom = "segment",y = -0.2,yend = ph_offset,x = cells_ml,xend = cells_ml,linetype = "dashed") +
  scale_x_log10() +
  scale_y_continuous(limits = c(-0.2,1.75),expand = expansion()) +
  annotate("text", y = 1.4, x = min(ph_cd$dfs[[sp[i]]]$CellDensity, na.rm = TRUE), size = 4, fontface = 'italic', 
         label = formal[sp[i]], hjust = 0, vjust = 1.1) +
  ggsci::scale_color_npg() +
  labs(x = "log10 Cell Density (cells mL-1)", y = "ΔpH",color = "Container",shape = "Air CO2 Level") +
  ggpubr::theme_pubclean() + theme(legend.position = "right")

}

cells_ml_ehux <- as.numeric(ph %>% subset(sp == "Ehux" & ph == ph_offset) %>% select(cd))
cells_ml_tpsu <- as.numeric(ph %>% subset(sp == "Tpsu" & ph == ph_offset) %>% select(cd))

pH <- ph_cd$dfs %>%
  bind_rows() %>%
  ggplot() +
  geom_point(aes(CellDensity,pH_offset,
                 color = Species,
                 ),size = 2) +
  geom_line(aes(cd,ph,color = "Ehux"),data = subset(ph,sp == "Ehux"),size = 1) +
  geom_line(aes(cd,ph,color = "Tpsu"),data = subset(ph,sp == "Tpsu"),size = 1) +
  annotate(geom = "segment",y = ph_offset,yend = ph_offset,x = 0,xend = cells_ml_ehux,linetype = "dashed") +
  annotate(geom = "segment",y = -0.2,yend = ph_offset,x = cells_ml_ehux,xend = cells_ml_ehux,linetype = "dashed") +
  annotate(geom = "segment",y = ph_offset,yend = ph_offset,x = 0,xend = cells_ml_tpsu,linetype = "dashed") +
  annotate(geom = "segment",y = -0.2,yend = ph_offset,x = cells_ml_tpsu,xend = cells_ml_tpsu,linetype = "dashed") +
  scale_x_log10(breaks = 10^(3:6)) +
  scale_y_continuous(limits = c(-0.2,1.75),expand = expansion()) +
  scale_color_manual(values = c('darkgreen','#5cd65c')) +
  labs(x = "Cell Density (log10 cells/mL)", y = "ΔpH",shape = "Air CO2 Level") +
  ggpubr::theme_pubclean() + theme(legend.position = "top")


#### Condition Panel
choose_biomes <- c("NA_STPS","NA_STSS","Atl_EQU")
#choose_biomes <- unique(abs$biome)
probs <- c(0.95,0.80,0.5)

vars <- colnames(lab)[str_detect(colnames(lab),"pyco2sys_")] %>%
  str_remove_all("_mean") %>% str_remove_all("_sd") %>% unique()

# lab_plot <- lab %>%
#   select(-pyco2sys_par1_mean) %>%
#   filter(samp_group %in% c("all_sample","all_blank")) %>%
#   pivot_longer(
#     cols = c(matches("^pyco2sys_.*_(mean|sd)$"), -starts_with("seacarb_flag_")),
#     names_to = c("var", ".value"),
#     names_pattern = "^(pyco2sys_.*)_(mean|sd)$"
#   ) %>%
#   mutate(value = set_errors(mean, sd)) %>%
#   select(-mean, -sd) %>%
#   pivot_wider(names_from = var, values_from = value) %>%
#   group_by(Incubator.CO2.ppm.nominal, Name) %>%
#   select(Incubator.CO2.ppm.nominal, Name, all_of(vars)) %>%
#   group_by(Incubator.CO2.ppm.nominal, Name) %>%
#   summarise(across(all_of(vars),.fns = list(mean = mean))) %>%
#   mutate(across(all_of(paste0(vars,"_mean")),.fns = list(sd = errors))) %>%
#   rename_with(~ gsub("_mean_sd", "_sd", .x))

lab_plot <- lab %>%
  filter(samp_group %in% c("all_sample"))
  

# Core refactor function
refactor_vars <- function(df, dataset, x_var, y_var, ymult = 1,xmult = 1, keep = NULL) {
  x_mean <- sym(paste0(x_var, "_mean"))
  x_sd   <- sym(paste0(x_var, "_sd"))
  y_mean <- sym(paste0(y_var, "_mean"))
  y_sd   <- sym(paste0(y_var, "_sd"))
  
  df %>%
    transmute(
      dataset = dataset,
      x    = !!x_mean * xmult,
      y    = !!y_mean * ymult,
      x_sd = !!x_sd   * xmult,
      y_sd = !!y_sd   * ymult,
      !!!syms(keep)
    )
}

# Wrapper: multiple datasets → bind_rows
refactor_many <- function(specs) {
  map_dfr(specs, function(s) {
    refactor_vars(
      df      = s$df,
      dataset = s$dataset,
      x_var   = s$x_var,
      y_var   = s$y_var,
      xmult    = s$xmult %||% 1,
      ymult    = s$ymult %||% 1,
      keep    = s$keep %||% NULL
    )
  })
}


# CMIP6 pre-processing step
cmip6_wide <- delta %>%
  subset(biome %in% choose_biomes) 

cmip6_wide$p_s <-  paste(cmip6_wide$scenario, cmip6_wide$period,  sep = " - ")

cmip6_wide[which(cmip6_wide$p_s %in% c("ssp245 - 1850s","ssp585 - 1850s")),"p_s"] <- "Historical , 1850s"
cmip6_wide[which(cmip6_wide$p_s %in% c("ssp245 - 1850s","ssp585 - 1850s")),"p_s"] <- "Historical , 1850s"
cmip6_wide[which(cmip6_wide$p_s %in% c("ssp585 - 2090s")),"p_s"] <- "SSP5-8.5 , 2090s"
cmip6_wide[which(cmip6_wide$p_s %in% c("ssp245 - 2090s")),"p_s"] <- "SSP2-4.5 , 2090s"
cmip6_wide[which(cmip6_wide$p_s %in% c("ssp585 - 2000s")),"p_s"] <- "Both Scenarios, 2000s"
cmip6_wide[which(cmip6_wide$p_s %in% c("ssp245 - 2000s")),"p_s"] <- "Both Scenarios, 2000s"

levels_order <- c(
  "Historical , 1850s",
  "Both Scenarios, 2000s",
  "SSP2-4.5 , 2090s",
  "SSP5-8.5 , 2090s",
  "BATS", "Lab Data, this study"
)
cmip6_wide$p_s <- factor(cmip6_wide$p_s,levels = levels_order)

cmip6_wide <- cmip6_wide %>% rename(aqueous_mean = "areaW_mean_delta_aqueous",
                      aqueous_sd = "sd_delta_aqueous",
                      bicarbonate_mean = "areaW_mean_delta_bicarbonate",
                      bicarbonate_sd = "sd_delta_bicarbonate",
                      carbonate_mean = "areaW_mean_delta_carbonate",
                      carbonate_sd = "sd_delta_carbonate",
                      tos_mean = "areaW_mean_delta_tos",
                      tos_sd = "sd_delta_tos")

# Define the list of datasets + specs
dat1 <- list(
  list(df = subset(BATS$means, nominal_depth == 0 & decade != "2020s"), dataset = "BATS", y_var = "CO2", x_var = "HCO3", ymult = 1, xmult = 1),
  list(df = cmip6_wide, dataset = "CMIP6", y_var = "aqueous", x_var = "bicarbonate", keep = c("p_s","type")),
  list(df = lab_plot, dataset = "Lab Data, this study", y_var = "pyco2sys_CO2", x_var = "pyco2sys_HCO3",
       ymult = 1e6,xmult = 1e6, keep = c("Incubator.CO2.ppm.nominal"))
) %>%
  refactor_many() %>%
  mutate(#samp_group = if_else(is.na(samp_group), "", samp_group),
    p_s = if_else(is.na(p_s), dataset, p_s),
    p_s = factor(p_s,levels_order),
    shape = type,
    shape = if_else(is.na(shape),dataset,shape))

errors(dat1$x) <- dat1$x_sd
errors(dat1$y) <- dat1$y_sd

baseline_lab_x <- dat1[which(dat1$Incubator.CO2.ppm.nominal == 420),] %>% pull(x)
baseline_lab_y <- dat1[which(dat1$Incubator.CO2.ppm.nominal == 420),] %>% pull(y)

baseline_BATS_x <- dat1[which(dat1$decade == "2000s"),] %>% pull(x)
baseline_BATS_y <- dat1[which(dat1$decade == "2000s"),] %>% pull(y)


dat1 <- dat1 %>%
  ungroup() %>%
  mutate(
    x = if_else(
      dataset == "Lab Data, this study",
      x - baseline_lab_x,
      x
    ),
    y = if_else(
      dataset == "Lab Data, this study",
      y - baseline_lab_y,
      y
    ),
    x = if_else(
      dataset == "BATS",
      x - baseline_BATS_x,
      x
    ),
    y = if_else(
      dataset == "BATS",
      y - baseline_BATS_y,
      y
    )
  )

dat1$x_sd <- errors(dat1$x) 
dat1$y_sd <- errors(dat1$y)

A <- dat1 %>% subset(x != 0 & dataset != "BATS") %>%
  ggplot(aes(x, y)) +
  geom_point(aes(color = p_s, shape = shape), size = 2, na.rm = TRUE) +
  geom_errorbar(aes(ymin = y - y_sd, ymax = y + y_sd, color = p_s), width = 0) +
  geom_errorbar(aes(xmin = x - x_sd, xmax = x + x_sd, color = p_s), width = 0) +
  labs(x = expression("Δ [HCO"[3]*"] (umol kg"^{-1}*")"),
       y = expression("Δ [CO"[2]*"] (umol kg"^{-1}*")")) +
  scale_color_manual(values = c("grey50","orange","maroon","darkgreen")) +
  scale_shape_manual(values = c("STPS" = 19,"STSS" = 17, "EQU" = 15,"Lab Data, this study" = 22, "BATS" = 10)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_line(aes(x, y,color = "BATS"),data = subset(dat1,dataset == "BATS"),color = "blue") +
  geom_point(aes(x, y,color = "BATS"),data = subset(dat1,dataset == "BATS"),color = "blue") +
theme(legend.position = "none") 
A

#####

# Define the list of datasets + specs
dat2 <- list(
  list(df = subset(BATS$means, nominal_depth == 0 & decade != "2020s"), dataset = "BATS", y_var = "CO2", x_var = "Temp", ymult = 1, xmult = 1),
  list(df = cmip6_wide, dataset = "CMIP6", y_var = "aqueous", x_var = "tos", keep = c("p_s","type")),
  list(df = lab_plot, dataset = "Lab Data, this study", y_var = "pyco2sys_CO2", x_var = "Temp.nominal",
       ymult = 1e6,xmult = 1, keep = c("Incubator.CO2.ppm.nominal"))
) %>%
  refactor_many() %>%
  mutate(#samp_group = if_else(is.na(samp_group), "", samp_group),
    p_s = if_else(is.na(p_s), dataset, p_s),
    p_s = factor(p_s,levels_order),
    shape = type,
    shape = if_else(is.na(shape),dataset,shape))

errors(dat2$x) <- dat2$x_sd
errors(dat2$y) <- dat2$y_sd

baseline_lab_x <- dat2[which(dat1$Incubator.CO2.ppm.nominal == 420),] %>% pull(x)
baseline_lab_y <- dat2[which(dat1$Incubator.CO2.ppm.nominal == 420),] %>% pull(y)

baseline_BATS_x <- dat2[which(dat1$decade == "2000s"),] %>% pull(x)
baseline_BATS_y <- dat2[which(dat1$decade == "2000s"),] %>% pull(y)


dat2 <- dat2 %>%
  mutate(
    x = if_else(
      dataset == "Lab Data, this study",
      x - baseline_lab_x,
      x
    ),
    y = if_else(
      dataset == "Lab Data, this study",
      y - baseline_lab_y,
      y
    ),
    x = if_else(
      dataset == "BATS",
      x - baseline_BATS_x,
      x
    ),
    y = if_else(
      dataset == "BATS",
      y - baseline_BATS_y,
      y
    )
  )

dat2$x_sd <- errors(dat2$x) 
dat2$y_sd <- errors(dat2$y)

B <- dat2 %>% subset(x != 0 & dataset != "BATS") %>%
  ggplot(aes(x, y)) +
  geom_point(aes(color = p_s, shape = shape), size = 2, na.rm = TRUE) +
  geom_errorbar(aes(ymin = y - y_sd, ymax = y + y_sd, color = p_s), width = 0) +
  geom_errorbar(aes(xmin = x - x_sd, xmax = x + x_sd, color = p_s), width = 0) +
  labs(x = expression("Δ Temperature (C)"),
       y = expression("Δ [CO"[2]*"] (umol kg"^{-1}*")")) +
  scale_color_manual(values = c("grey50","orange","maroon","darkgreen")) +
  scale_shape_manual(values = c("STPS" = 19,"STSS" = 17, "EQU" = 15,"Lab Data, this study" = 22, "BATS" = 10)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_line(aes(x, y),data = subset(dat2,dataset == "BATS"),color = "blue") +
  geom_point(aes(x, y),data = subset(dat2,dataset == "BATS"),color = "blue") +
  theme(legend.position = "none") 

A + theme_minimal() + B + theme_minimal() + plot_layout(guides = "collect")

A + B + A + B +
plot_layout(guides = "collect",nrow = 2) 

# Define the list of datasets + specs
dat3 <- list(
  list(df = subset(BATS$means, nominal_depth == 0 & decade != "2020s"), dataset = "BATS", y_var = "CO2", x_var = "CO3", ymult = 1e6, xmult = 1e6),
  list(df = cmip6_wide, dataset = "CMIP6", y_var = "aqueous", x_var = "carbonate", keep = c("p_s","type")),
  list(df = lab_plot, dataset = "Lab Data, this study", y_var = "pyco2sys_CO2", x_var = "pyco2sys_CO3",
       ymult = 1e6,xmult = 1e6, keep = c("Incubator.CO2.ppm.nominal"))
) %>%
  refactor_many() %>%
  mutate(#samp_group = if_else(is.na(samp_group), "", samp_group),
    p_s = if_else(is.na(p_s), dataset, p_s),
    p_s = factor(p_s,levels_order),
    shape = type,
    shape = if_else(is.na(shape),dataset,shape))

errors(dat3$x) <- dat3$x_sd
errors(dat3$y) <- dat3$y_sd

baseline_lab_x <- dat3[which(dat3$Incubator.CO2.ppm.nominal == 420),] %>% pull(x)
baseline_lab_y <- dat3[which(dat3$Incubator.CO2.ppm.nominal == 420),] %>% pull(y)

baseline_BATS_x <- dat3[which(dat3$decade == "2000s"),] %>% pull(x)
baseline_BATS_y <- dat3[which(dat3$decade == "2000s"),] %>% pull(y)


dat3 <- dat3 %>%
  ungroup() %>%
  mutate(
    x = if_else(
      dataset == "Lab Data, this study",
      x - baseline_lab_x,
      x
    ),
    y = if_else(
      dataset == "Lab Data, this study",
      y - baseline_lab_y,
      y
    ),
    x = if_else(
      dataset == "BATS",
      x - baseline_BATS_x,
      x
    ),
    y = if_else(
      dataset == "BATS",
      y - baseline_BATS_y,
      y
    )
  )

dat3$x_sd <- errors(dat3$x) 
dat3$y_sd <- errors(dat3$y)

C <- dat3 %>% subset(x != 0 & dataset != "BATS") %>%
  ggplot(aes(x, y)) +
  geom_point(aes(color = p_s, shape = shape), size = 2, na.rm = TRUE) +
  geom_errorbar(aes(ymin = y - y_sd, ymax = y + y_sd, color = p_s), width = 0) +
  geom_errorbar(aes(xmin = x - x_sd, xmax = x + x_sd, color = p_s), width = 0) +
  labs(x = expression("Δ [CO"[3]*"] (umol kg"^{-1}*")"),
       y = expression("Δ [CO"[2]*"] (umol kg"^{-1}*")")) +
  scale_color_manual(values = c("grey50","orange","maroon","darkgreen")) +
  scale_shape_manual(values = c("STPS" = 19,"STSS" = 17, "EQU" = 15,"Lab Data, this study" = 22, "BATS" = 10)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_line(aes(x, y),data = subset(dat1,dataset == "BATS"),color = "blue") +
  geom_point(aes(x, y),data = subset(dat1,dataset == "BATS"),color = "blue") +
  theme(legend.position = "none")

C

#####
pretty_sci <- function(x) {
  parse(text = gsub("e\\+?", " %*% 10^", format(x, scientific = TRUE)))
}



  pH + theme(axis.title = element_text(size = 10)) + theme(legend.position = "none") +
  plot_spacer() +
  #ehux_acclimation + theme(axis.title = element_text(size = 10),axis.text.y.left = element_blank()) +
  A + theme(axis.title = element_text(size = 10)) +
  B + theme(axis.title = element_text(size = 10)) + theme(legend.position = "none") +
  plot_layout(ncol = 2, axes = "collect",guides = "collect") +
  #plot_annotation(tag_levels = c("a")) &
  theme(
    plot.tag = element_text(
      face = "bold",
      family = "Helvetica", # or "Arial"
      size = 14,
      hjust = -0.1,
      vjust = 1.1
    )
  )

#ggsave("C:/Users/henryholm/OneDrive - Woods Hole Oceanographic Institution/Desktop/Galen_CMIP_Collab/Figures/Fig2_condition_update.pdf",
#       device = cairo_pdf, width = 5.5, height = 8)

A + B + theme(legend.position = "top")



ggsave("C:/Users/henryholm/OneDrive - Woods Hole Oceanographic Institution/Desktop/Galen_CMIP_Collab/Figures/Fig2_condition_final.pdf",
       device = cairo_pdf, width = 5.5, height = 6)

A + theme(legend.position = "left")

ggsave("C:/Users/henryholm/OneDrive - Woods Hole Oceanographic Institution/Desktop/Galen_CMIP_Collab/Figures/Fig3_legend.pdf",
       device = cairo_pdf, width = 5.5, height = 6)

############### abs version for suppliment

# CMIP6 pre-processing step
cmip6_wide <- delta %>%
  subset(biome %in% choose_biomes) 

cmip6_wide$p_s <-  paste(cmip6_wide$scenario, cmip6_wide$period,  sep = " - ")

cmip6_wide[which(cmip6_wide$p_s %in% c("ssp245 - 1850s","ssp585 - 1850s")),"p_s"] <- "Historical , 1850s"
cmip6_wide[which(cmip6_wide$p_s %in% c("ssp245 - 1850s","ssp585 - 1850s")),"p_s"] <- "Historical , 1850s"
cmip6_wide[which(cmip6_wide$p_s %in% c("ssp585 - 2090s")),"p_s"] <- "SSP5-8.5 , 2090s"
cmip6_wide[which(cmip6_wide$p_s %in% c("ssp245 - 2090s")),"p_s"] <- "SSP2-4.5 , 2090s"
cmip6_wide[which(cmip6_wide$p_s %in% c("ssp585 - 2000s")),"p_s"] <- "Both Scenarios, 2000s"
cmip6_wide[which(cmip6_wide$p_s %in% c("ssp245 - 2000s")),"p_s"] <- "Both Scenarios, 2000s"

levels_order <- c(
  "Historical , 1850s",
  "Both Scenarios, 2000s",
  "SSP2-4.5 , 2090s",
  "SSP5-8.5 , 2090s",
  "BATS", "Lab Data, this study"
)
cmip6_wide$p_s <- factor(cmip6_wide$p_s,levels = levels_order)

cmip6_wide <- cmip6_wide %>% rename(aqueous_mean = "areaW_mean_abs_aqueous",
                      aqueous_sd = "sd_abs_aqueous",
                      bicarbonate_mean = "areaW_mean_abs_bicarbonate",
                      bicarbonate_sd = "sd_abs_bicarbonate",
                      carbonate_mean = "areaW_mean_abs_carbonate",
                      carbonate_sd = "sd_abs_carbonate",
                      pH_mean = "areaW_mean_abs_pH",
                      pH_sd = "sd_abs_pH",
                      tos_mean = "areaW_mean_abs_tos",
                      tos_sd = "sd_abs_tos")
# Define the list of datasets + specs
dat1 <- list(
  list(df = subset(BATS$means, nominal_depth == 0 & decade != "2020s"), dataset = "BATS", y_var = "CO2", x_var = "HCO3", ymult = 1, xmult = 1),
  list(df = cmip6_wide, dataset = "CMIP6", y_var = "aqueous", x_var = "bicarbonate", keep = c("p_s","type")),
  list(df = lab_plot, dataset = "Lab Data, this study", y_var = "pyco2sys_CO2", x_var = "pyco2sys_HCO3",
       ymult = 1e6,xmult = 1e6, keep = c("Incubator.CO2.ppm.nominal"))
) %>%
  refactor_many() %>%
  mutate(#samp_group = if_else(is.na(samp_group), "", samp_group),
    p_s = if_else(is.na(p_s), dataset, p_s),
    p_s = factor(p_s,levels_order),
    shape = type,
    shape = if_else(is.na(shape),dataset,shape))


A <- dat1 %>% # subset(dataset != "BATS") %>%
  ggplot(aes(x, y)) +
  geom_point(aes(color = p_s, shape = shape), size = 2, na.rm = TRUE) +
  geom_errorbar(aes(ymin = y - y_sd*2, ymax = y + y_sd*2, color = p_s), width = 0) +
  geom_errorbar(aes(xmin = x - x_sd*2, xmax = x + x_sd*2, color = p_s), width = 0) +
  labs(x = expression("[HCO"[3]*"] (umol kg"^{-1}*")"),
       y = expression("[CO"[2]*"] (umol kg"^{-1}*")")) +
  scale_color_manual(values = c("grey50","#3295a8","orange","maroon","blue","darkgreen")) +
  scale_shape_manual(values = c("STPS" = 19,"STSS" = 17,"EQU" = 15, "Lab Data, this study" = 22,"BATS" = 2)) +
  theme(legend.position = "none") 

A

# Define the list of datasets + specs
dat2 <- list(
  list(df = subset(BATS$means, nominal_depth == 0 & decade != "2020s"), dataset = "BATS", y_var = "CO2", x_var = "Temp", ymult = 1, xmult = 1),
  list(df = cmip6_wide, dataset = "CMIP6", y_var = "aqueous", x_var = "tos", keep = c("p_s","type")),
  list(df = lab_plot, dataset = "Lab Data, this study", y_var = "pyco2sys_CO2", x_var = "Temp.nominal",
       ymult = 1e6,xmult = 1, keep = c("Incubator.CO2.ppm.nominal"))
) %>%
  refactor_many() %>%
  mutate(#samp_group = if_else(is.na(samp_group), "", samp_group),
    p_s = if_else(is.na(p_s), dataset, p_s),
    p_s = factor(p_s,levels_order),
    shape = type,
    shape = if_else(is.na(shape),dataset,shape))

B <- dat2 %>% #subset(x != 0 & dataset != "BATS") %>%
  ggplot(aes(x, y)) +
  geom_point(aes(color = p_s, shape = shape), size = 2, na.rm = TRUE) +
  geom_errorbar(aes(ymin = y - y_sd*2, ymax = y + y_sd*2, color = p_s), width = 0) +
  geom_errorbar(aes(xmin = x - x_sd*2, xmax = x + x_sd*2, color = p_s), width = 0) +
  labs(x = expression("Temperature (C)"),
       y = expression("[CO"[2]*"] (umol kg"^{-1}*")")) +
  scale_color_manual(values = c("grey50","#3295a8","orange","maroon","blue","darkgreen")) +
  scale_shape_manual(values = c("STPS" = 19,"STSS" = 17,"EQU" = 15, "Lab Data, this study" = 22,"BATS" = 2)) +
  geom_line(aes(x, y),data = subset(dat2,dataset == "BATS"),color = "blue") +
  geom_point(aes(x, y),data = subset(dat2,dataset == "BATS"),color = "blue") +
  theme(legend.position = "none") 

A + B

B + theme(legend.position = "right")

# Define the list of datasets + specs
dat3 <- list(
  list(df = subset(BATS$means, nominal_depth == 0 & decade != "2020s"), dataset = "BATS", y_var = "CO2", x_var = "CO3", ymult = 1, xmult = 1),
  list(df = cmip6_wide, dataset = "CMIP6", y_var = "aqueous", x_var = "carbonate", keep = c("p_s","type")),
  list(df = lab_plot, dataset = "Lab Data, this study", y_var = "pyco2sys_CO2", x_var = "pyco2sys_CO3",
       ymult = 1e6,xmult = 1e6, keep = c("Incubator.CO2.ppm.nominal"))
) %>%
  refactor_many() %>%
  mutate(#samp_group = if_else(is.na(samp_group), "", samp_group),
    p_s = if_else(is.na(p_s), dataset, p_s),
    p_s = factor(p_s,levels_order),
    shape = type,
    shape = if_else(is.na(shape),dataset,shape))

C <- dat3 %>% # subset(dataset != "BATS") %>%
  ggplot(aes(x, y)) +
  geom_point(aes(color = p_s, shape = shape), size = 2, na.rm = TRUE) +
  geom_errorbar(aes(ymin = y - y_sd*2, ymax = y + y_sd*2, color = p_s), width = 0) +
  geom_errorbar(aes(xmin = x - x_sd*2, xmax = x + x_sd*2, color = p_s), width = 0) +
  labs(x = expression("[CO"[3]*"] (umol kg"^{-1}*")"),
       y = expression("[CO"[2]*"] (umol kg"^{-1}*")")) +
  scale_color_manual(values = c("grey50","#3295a8","orange","maroon","blue","darkgreen")) +
  scale_shape_manual(values = c("STPS" = 19,"STSS" = 17,"EQU" = 15, "Lab Data, this study" = 22,"BATS" = 2)) +
  theme(legend.position = "none") 


# Define the list of datasets + specs
dat4 <- list(
  list(df = subset(BATS$means, nominal_depth == 0 & decade != "2020s"), dataset = "BATS", y_var = "CO2", x_var = "pH", ymult = 1, xmult = 1),
  list(df = cmip6_wide, dataset = "CMIP6", y_var = "aqueous", x_var = "pH", keep = c("p_s","type")),
  list(df = lab_plot, dataset = "Lab Data, this study", y_var = "pyco2sys_CO2", x_var = "pyco2sys_pH_total",
       ymult = 1e6,xmult = 1, keep = c("Incubator.CO2.ppm.nominal"))
) %>%
  refactor_many() %>%
  mutate(#samp_group = if_else(is.na(samp_group), "", samp_group),
    p_s = if_else(is.na(p_s), dataset, p_s),
    p_s = factor(p_s,levels_order),
    shape = type,
    shape = if_else(is.na(shape),dataset,shape))

D <- dat4 %>% # subset(dataset != "BATS") %>%
  ggplot(aes(x, y)) +
  geom_point(aes(color = p_s, shape = shape), size = 2, na.rm = TRUE) +
  geom_errorbar(aes(ymin = y - y_sd*2, ymax = y + y_sd*2, color = p_s), width = 0) +
  geom_errorbar(aes(xmin = x - x_sd*2, xmax = x + x_sd*2, color = p_s), width = 0) +
  labs(x = "pH (Total Scale)",
       y = expression("[CO"[2]*"] (umol kg"^{-1}*")")) +
  scale_color_manual(values = c("grey50","#3295a8","orange","maroon","blue","darkgreen")) +
  scale_shape_manual(values = c("STPS" = 19,"STSS" = 17,"EQU" = 15, "Lab Data, this study" = 22,"BATS" = 2)) +
  theme(legend.position = "none") 

A + B + C + D

check_points(co2 = lab_plot$pyco2sys_CO2_mean*1e6,
             hco3 = lab_plot$pyco2sys_HCO3_mean*1e6,
             co3 = lab_plot$pyco2sys_CO3_mean*1e6,
             pH = lab_plot$pyco2sys_pH_total_mean,
             lab_plot$Temp.nominal_mean,carbon_regions = carbon_regions
             )

    sd_vec <- c(
      sd(noice$areaW_mean_abs_aqueous),
      sd(noice$areaW_mean_abs_bicarbonate),
      sd(noice$areaW_mean_abs_carbonate),
      sd(noice$areaW_mean_abs_tos),
      sd(noice$areaW_mean_abs_pH)
    )

    d <- sqrt(
      ((lab_plot$pyco2sys_CO2_mean[1]*1e6 - noice$areaW_mean_abs_aqueous) / sd_vec[1])^2 +
        ((lab_plot$pyco2sys_HCO3_mean[1]*1e6 - noice$areaW_mean_abs_bicarbonate) / sd_vec[2])^2 +
        ((lab_plot$pyco2sys_CO3_mean[1]*1e6 - noice$areaW_mean_abs_carbonate) / sd_vec[3])^2 +
        ((lab_plot$Temp.nominal_mean[1] - noice$areaW_mean_abs_tos) / sd_vec[4])^2 +
        ((lab_plot$pyco2sys_pH_total_mean[1] - noice$areaW_mean_abs_pH) / sd_vec[5])^2
    )
    closest_idx <- which.min(d)
    noice[closest_idx,c("scenario","biome","period")]
    
    d <- sqrt(
      ((lab_plot$pyco2sys_CO2_mean[2]*1e6 - noice$areaW_mean_abs_aqueous) / sd_vec[1])^2 +
        ((lab_plot$pyco2sys_HCO3_mean[2]*1e6 - noice$areaW_mean_abs_bicarbonate) / sd_vec[2])^2 +
        ((lab_plot$pyco2sys_CO3_mean[2]*1e6 - noice$areaW_mean_abs_carbonate) / sd_vec[3])^2 +
        ((lab_plot$Temp.nominal_mean[2] - noice$areaW_mean_abs_tos) / sd_vec[4])^2 +
        ((lab_plot$pyco2sys_pH_total_mean[2] - noice$areaW_mean_abs_pH) / sd_vec[5])^2
    )
    closest_idx <- which.min(d)
    noice[closest_idx,c("scenario","biome","period")]
    noice[order(d),c("scenario","biome","period")]
    
lab_plot$pyco2sys_CO2_mean

lab_plot$pyco2sys_CO2_sd*1e6
lab_plot$pyco2sys_HCO3_sd*1e6
lab_plot$pyco2sys_CO3_sd*1e6
lab_plot$pyco2sys_pH_total_sd

ggplot() + geom_histogram(aes(noice$sd_abs_aqueous)) + geom_vline()

OA_ICC %>% subset(CO2_sd != 0) %>% ggplot() + geom_density(aes(CO2_sd))

range(noice$sd_abs_bicarbonate)
range(noice$sd_abs_carbonate)
range(noice$sd_abs_pH)

```

```{r}

pyco2$sys()
```

Figure 4 - Compare to OA-ICC
```{r}
#write.csv(delta,"C:/Users/henryholm/OneDrive - Woods Hole Oceanographic Institution/delta.csv")

delta <- read.csv("C:/Users/henryholm/OneDrive - Woods Hole Oceanographic Institution/delta.csv") %>% as_tibble()
OA_ICC <- read.csv("C:/Users/henryholm/OneDrive - Woods Hole Oceanographic Institution/Desktop/Galen_CMIP_Collab/carb_data_merged.csv")

length(unique(OA_ICC$Study_DOI))
nrow(OA_ICC)

OA_ICC <- OA_ICC %>% subset(!(is.na(CO2_mean) | is.na(HCO3_mean) | is.na(CO3_mean)))

length(unique(OA_ICC$Study_DOI))
nrow(OA_ICC)

dhco3 <- expression(paste("Δ [HCO"[3]^"-","]", " (", "umol ", kg^{-1}, ")"))
dco2 <- expression("Δ [CO"[2]*"] (umol kg"^{-1}*")")
dco3 <- expression(paste("Δ [CO"[3]^"-2","]", " (", "umol ", kg^{-1}, ")"))
dtemp <- "Δ Temperature (°C)"
dpco2 <- expression("Δ pCO"[2]*"")

hco3 <- expression(paste("[HCO"[3]^"-","]", " (", "umol ", kg^{-1}, ")"))
co2 <- expression("[CO"[2]*"] (umol kg"^{-1}*")")
co3 <- expression(paste("[CO"[3]^"-2","]", " (", "umol ", kg^{-1}, ")"))
temp <- "Temperature (°C)"
pco2 <- expression("pCO"[2]*"")

inset <- ggplot(
    delta %>% subset(type != "ICE" & period != "2000s"),
    aes(x = areaW_mean_delta_aqueous, y = areaW_mean_delta_bicarbonate, color = p_s)
  ) +
    #ggdensity::geom_hdr(aes(CO2_delta, HCO3_delta,color = NA),
    #                  fill = 'grey70',data = OA_ICC) +
    geom_point(aes(x = CO2_delta, y = HCO3_delta, color = 'Lit data'),
            alpha = 0.3, data = OA_ICC,shape = 19,color = "grey70") +
    geom_point(aes(shape = type),size = 2) +
    geom_errorbar(aes(
      ymin = areaW_mean_delta_bicarbonate- sd_delta_bicarbonate,
      ymax = areaW_mean_delta_bicarbonate + sd_delta_bicarbonate
    ),linewidth = 0.1) +
    geom_errorbarh(aes(
      xmin = areaW_mean_delta_aqueous - sd_delta_aqueous,
      xmax = areaW_mean_delta_aqueous + sd_delta_aqueous
    ),linewidth = 0.1) +
    labs(
      y = dhco3,
      x = dco2,
      color = "Period and Scenario", shape = "Biome Type"
    ) +
    scale_color_manual(values = c("grey30", "orange", "maroon")) +
    scale_shape_manual(values = c(15:17,4)) +
  theme_minimal()

main <- inset +
  coord_cartesian(xlim = c(-25,100),c(-100,1100)) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.1, linetype = "dashed") +
  geom_vline(xintercept = 0, color = "black", linewidth = 0.1, linetype = "dashed")

panel1 <- main  + theme(legend.position = "none") + patchwork::inset_element(
  inset +
    theme(
      axis.title = element_blank(), # remove axis labels
      legend.position = "none", # remove legend
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background = element_rect(fill = "white", colour = "black")
    ),
  left = 0.45, bottom = 0.45,
  right = 0.95, top = 0.95
)

inset <- ggplot(
    delta %>% subset(type != "ICE" & period != "2000s"),
    aes(x = areaW_mean_delta_aqueous, y = areaW_mean_delta_carbonate, color = p_s)
  ) +
    #ggdensity::geom_hdr(aes(CO2_delta, CO3_delta,color = NA),
   #                   fill = 'grey70',data = OA_ICC) +
  geom_point(aes(x = CO2_delta, y = CO3_delta, color = 'Lit data'),
            alpha = 0.3, data = OA_ICC,shape = 19,color = "grey70") +
    geom_point(aes(shape = type),size = 2) +
    geom_errorbar(aes(
      ymin = areaW_mean_delta_carbonate- sd_delta_carbonate,
      ymax = areaW_mean_delta_carbonate + sd_delta_carbonate
    ),linewidth = 0.1) +
    geom_errorbarh(aes(
      xmin = areaW_mean_delta_aqueous - sd_delta_aqueous,
      xmax = areaW_mean_delta_aqueous + sd_delta_aqueous
    ),linewidth = 0.1) +
    labs(
      y = dco3,
      x = dco2,
      color = "Period and Scenario", shape = "Biome Type"
    ) +
    scale_color_manual(values = c("grey30", "orange", "maroon")) +
    scale_shape_manual(values = c(15:17,4)) +
  theme_minimal()

main <- inset +
  coord_cartesian(xlim = c(-25,100),c(-300,500)) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.1, linetype = "dashed") +
  geom_vline(xintercept = 0, color = "black", linewidth = 0.1, linetype = "dashed")

panel2 <- main  + theme(legend.position = "none") + inset_element(
  inset +
    theme(
      axis.title = element_blank(), # remove axis labels
      legend.position = "none", # remove legend
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background = element_rect(fill = "white", colour = "black")
    ),
  left = 0.45, bottom = 0.45,
  right = 0.95, top = 0.95
)

panel1 + panel2

#ggsave("C:/Users/henryholm/OneDrive - Woods Hole Oceanographic Institution/Desktop/Galen_CMIP_Collab/Figures/Fig4_delta.pdf",
#       device = cairo_pdf, width = 8.5, height = 4)

inset <- delta %>%
  subset(type != "ICE" & p_s != "ssp585 - 2000s") %>%
  group_by(scenario, biome, loc, type, period) %>%
  ungroup() %>%
  ggplot(aes(y = areaW_mean_abs_bicarbonate, x = areaW_mean_abs_aqueous, color = p_s)) +
  #ggdensity::geom_hdr(aes(y = HCO3_mean, x = CO2_mean, color = NA),
  #                fill = 'grey70',data = OA_ICC) +
  geom_point(aes(y = HCO3_mean, x = CO2_mean, color = NA),
                  color = 'grey70',alpha = 0.3,shape = 19, data = OA_ICC) +
  geom_point(aes(shape = type)) +
  ggdensity::geom_hdr_points(aes(y = HCO3_mean, x = CO2_mean, color = NA),
                  color = 'grey70',alpha = 0.3,shape = 19, data = OA_ICC) +
  geom_errorbar(aes(xmin = areaW_mean_abs_aqueous - sd_abs_aqueous*2,
                    xmax = areaW_mean_abs_aqueous + sd_abs_aqueous*2),linewidth = 0.1) +
  geom_errorbar(aes(ymin = areaW_mean_abs_bicarbonate - sd_abs_bicarbonate*2,
                    ymax = areaW_mean_abs_bicarbonate + sd_abs_bicarbonate*2),linewidth = 0.1) +
  labs(x = co2, y = hco3,color = "Period and Scenario",shape = "Biome Type") +
  scale_color_manual(values = c("grey20","orange","#3295a8","maroon")) +
  plot_layout(ncol = 1,guides = "collect") +
  theme_minimal() +
  scale_shape_manual(values = list("EQU" = 15, "STPS" = 16, "STSS" = 17, "SPSS" = 4))

main <- inset + coord_cartesian(xlim = c(0,100),c(0,2600))

panel3 <- main +# theme(legend.position = "none") +
inset_element(
  inset +
    theme(
      axis.title = element_blank(), # remove axis labels
      legend.position = "none", # remove legend
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background = element_rect(fill = "white", colour = "black")
    ),
  left = 0.45, bottom = 0.05,
  right = 0.95, top = 0.45    
)

#lb <- dat1 %>% subset(dataset == "Lab, This Study")

#main +
#  geom_point(aes(y = x, x = y), size = 2, data = lb) +
#  geom_errorbar(aes(y = x, x = y,xmin = y - y_sd, xmax = y + y_sd), width = 0, data = lb) +
#  geom_errorbar(aes(y = x, x = y,ymin = x - x_sd, ymax = x + x_sd), width = 0,data = lb) 

inset <- delta %>%
  subset(type != "ICE" & p_s != "ssp585 - 2000s") %>%
  group_by(scenario, biome, loc, type, period) %>%
  ungroup() %>%
  ggplot(aes(y = areaW_mean_abs_carbonate, x = areaW_mean_abs_aqueous, color = p_s)) +
  #ggdensity::geom_hdr(aes(y = CO3_mean, x = CO2_mean, color = NA),
  #                fill = 'grey70',data = OA_ICC) +
      geom_point(aes(y = CO3_mean, x = CO2_mean, color = 'Lit data'),
            alpha = 0.3, data = OA_ICC,shape = 19,color = "grey70") +
  geom_point(aes(shape = type)) +
  geom_errorbar(aes(xmin = areaW_mean_abs_aqueous - sd_abs_aqueous*2,
                    xmax = areaW_mean_abs_aqueous + sd_abs_aqueous*2),linewidth = 0.1) +
  geom_errorbar(aes(ymin = areaW_mean_abs_carbonate - sd_abs_carbonate*2,
                    ymax = areaW_mean_abs_carbonate + sd_abs_carbonate*2),linewidth = 0.1) +
  labs(x = co2, y = co3,color = "Period and Scenario",shape = "Biome Type") +
  scale_color_manual(values = c("grey20","orange","#3295a8","maroon")) +
  plot_layout(ncol = 1,guides = "collect") +
  theme_minimal() +
  scale_shape_manual(values = c(15:17,4))

main <- inset + coord_cartesian(xlim = c(0,100),ylim = c(0,600))

panel4 <- main + theme(legend.position = "none") + inset_element(
  inset +
    theme(
      axis.title = element_blank(), # remove axis labels
      legend.position = "none", # remove legend
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background = element_rect(fill = "white", colour = "black")
    ),
  left = 0.45, bottom = 0.45,
  right = 0.95, top = 0.95
) + labs(y = hco3, x = co3)


inset <- delta %>%
  subset(type != "ICE" & p_s != "ssp585 - 2000s") %>%
  group_by(scenario, biome, loc, type, period) %>%
  ungroup() %>%
  ggplot(aes(y = areaW_mean_abs_pH, x = areaW_mean_abs_aqueous, color = p_s)) +
      geom_point(aes(y = pHT_mean, x = CO2_mean, color = 'Lit data'),
            alpha = 0.3, data = OA_ICC,shape = 19,color = "grey70") +
  geom_point(aes(shape = type)) +
  geom_errorbar(aes(ymin = areaW_mean_abs_pH - sd_abs_pH*2,
                    ymax = areaW_mean_abs_pH + sd_abs_pH*2),linewidth = 0.1) +
  geom_errorbar(aes(xmin = areaW_mean_abs_aqueous - sd_abs_aqueous*2,
                    xmax = areaW_mean_abs_aqueous + sd_abs_aqueous*2),linewidth = 0.1) +
  labs(x = co2, y = 'pH (Total Scale)',color = "Period and Scenario",shape = "Biome Type") +
  scale_color_manual(values = c("grey20","orange","#3295a8","maroon")) +
  plot_layout(ncol = 1, guides = "collect") +
  theme_minimal() +
  scale_shape_manual(values = c(15:17,4))

main <- inset + coord_cartesian(xlim = c(0,100),ylim = c(7.2,9.2))

panel5 <- main + theme(legend.position = "none") + inset_element(
  inset +
    theme(
      axis.title = element_blank(), # remove axis labels
      legend.position = "none", # remove legend
      panel.background = element_rect(fill = "white", colour = NA),
      plot.background = element_rect(fill = "white", colour = "black")
    ),
  left = 0.45, bottom = 0.45,
  right = 0.95, top = 0.95
) + labs(y = hco3, x = co3)

panel3 + panel4 + plot_layout(ncol = 2)

ggsave("C:/Users/henryholm/OneDrive - Woods Hole Oceanographic Institution/Desktop/Galen_CMIP_Collab/Figures/Fig2_abs_final.pdf",
       device = cairo_pdf, width = 10, height = 4)

panel3 + panel4 + panel5 + plot_layout(ncol = 3) + theme(legend.position = "none")

ggsave("C:/Users/henryholm/OneDrive - Woods Hole Oceanographic Institution/Desktop/Galen_CMIP_Collab/Figures/Fig2_abs_final_w_pH2.pdf",
       device = cairo_pdf, width = 14, height = 4)

oa_icc_hits <- list()

within_k_sd <- function(x, mean, sd, k = 2) {
  (x > (mean - (k * sd))) & (x < (mean + (k * sd)))
}

noice <- delta %>% subset(type != "ICE")

for (i in 1:nrow(OA_ICC)) {
  # ind hits
  co2_hits  <- within_k_sd(OA_ICC[i, "CO2_mean"], noice$areaW_mean_abs_aqueous, noice$sd_abs_aqueous) 
  hco3_hits <- within_k_sd(OA_ICC[i, "HCO3_mean"], noice$areaW_mean_abs_bicarbonate, noice$sd_abs_bicarbonate) 
  co3_hits <- within_k_sd(OA_ICC[i, "CO3_mean"], noice$areaW_mean_abs_carbonate, noice$sd_abs_carbonate)
  temp_hits <- within_k_sd(OA_ICC[i,"Temp_mean"], noice$areaW_mean_abs_tos, noice$sd_abs_tos)
  
    OA_ICC[i, "co2_rel"] <- any(co2_hits,na.rm = T)
    OA_ICC[i, "hco3_rel"] <- any(hco3_hits,na.rm = T)
    OA_ICC[i, "co3_rel"] <- any(co3_hits,na.rm = T)
    OA_ICC[i, "temp_rel"] <- any(temp_hits,na.rm = T)
  
  # determine if it is ocean relevant 
  total_hits <- within_k_sd(OA_ICC[i, "CO2_mean"], noice$areaW_mean_abs_aqueous, noice$sd_abs_aqueous) &
    within_k_sd(OA_ICC[i, "HCO3_mean"], noice$areaW_mean_abs_bicarbonate, noice$sd_abs_bicarbonate) &
    within_k_sd(OA_ICC[i, "CO3_mean"], noice$areaW_mean_abs_carbonate, noice$sd_abs_carbonate)

  cols <- c("scenario", "biome", "loc", "type", "period") # to assign
  if (any(total_hits,na.rm = T)) {
    
    # what point is it closest too - evaluate all points so there is no sd bias 
    d <- sqrt(
      ((OA_ICC[i, "CO2_mean"] - noice$areaW_mean_abs_aqueous))^2 +
        ((OA_ICC[i, "HCO3_mean"] - noice$areaW_mean_abs_bicarbonate))^2 +
        ((OA_ICC[i, "CO3_mean"] - noice$areaW_mean_abs_carbonate))^2 +
        ((OA_ICC[i, "Temp_mean"] - noice$areaW_mean_abs_tos))^2
    )

    closest_idx <- which.min(d)

    oa_icc_hits[[i]] <- noice[closest_idx, ]
    OA_ICC[i, "ocean_relevant"] <- TRUE
    OA_ICC[i, cols] <- noice[closest_idx, cols]
  } else {
    oa_icc_hits[[i]] <- NULL
    OA_ICC[i, "ocean_relevant"] <- FALSE
    OA_ICC[i, cols] <- "Non-relevant"
  }
}

sum(OA_ICC$ocean_relevant == TRUE)/length(OA_ICC$ocean_relevant)

for (i in 1:nrow(OA_ICC)) {
  # ind hits
  co2_hits  <- within_k_sd(OA_ICC[i, "CO2_delta"], noice$areaW_mean_delta_aqueous, noice$sd_delta_aqueous) 
  hco3_hits <- within_k_sd(OA_ICC[i, "HCO3_delta"], noice$areaW_mean_delta_bicarbonate, noice$sd_delta_bicarbonate) 
  co3_hits <- within_k_sd(OA_ICC[i, "CO3_delta"], noice$areaW_mean_delta_carbonate, noice$sd_delta_carbonate)
  temp_hits <- within_k_sd(OA_ICC[i,"Temp_delta"], noice$areaW_mean_delta_tos, noice$sd_delta_tos)
  
  OA_ICC[i, "co2_rel_delta"] <- any(co2_hits,na.rm = T)
  OA_ICC[i, "hco3_rel_delta"] <- any(hco3_hits,na.rm = T)
  OA_ICC[i, "co3_rel_delta"] <- any(co3_hits,na.rm = T)
  OA_ICC[i, "temp_rel_delta"] <- any(temp_hits,na.rm = T)
  
  # determine if it is ocean relevant 
  total_hits <- within_k_sd(OA_ICC[i, "CO2_delta"], noice$areaW_mean_delta_aqueous, noice$sd_delta_aqueous) &
    within_k_sd(OA_ICC[i, "HCO3_delta"], noice$areaW_mean_delta_bicarbonate, noice$sd_delta_bicarbonate) &
    within_k_sd(OA_ICC[i, "CO3_delta"], noice$areaW_mean_delta_carbonate, noice$sd_delta_carbonate)

  cols <- c("scenario", "biome", "loc", "type", "period") # to assign
  cols_delta <- paste0(cols,"_delta")
  
  if (any(total_hits,na.rm = T)) {
    
    # what point is it closest too - evaluate all points so there is no sd bias 
    d <- sqrt(
      ((OA_ICC[i, "CO2_delta"] - noice$areaW_mean_delta_aqueous))^2 +
        ((OA_ICC[i, "HCO3_delta"] - noice$areaW_mean_delta_bicarbonate))^2 +
        ((OA_ICC[i, "CO3_delta"] - noice$areaW_mean_delta_carbonate))^2 
    )

    closest_idx <- which.min(d)

    oa_icc_hits[[i]] <- noice[closest_idx, ]
    OA_ICC[i, "ocean_relevant_delta"] <- TRUE
    OA_ICC[i, cols_delta] <- noice[closest_idx, cols]
  } else {
    oa_icc_hits[[i]] <- NULL
    OA_ICC[i, "ocean_relevant_delta"] <- FALSE
    OA_ICC[i, cols_delta] <- "Non-relevant"
  }
}

sum(OA_ICC$ocean_relevant_delta == TRUE)/length(OA_ICC$ocean_relevant_delta)

rel_doi <- OA_ICC %>%
  group_by(Study_DOI) %>%
  summarize(rel = any(ocean_relevant),
            rel_delta = any(ocean_relevant_delta),
            rel_delta_co2 = any(co2_rel_delta),
            ssp585_delta = any(scenario_delta == "ssp585"),
            ssp245_delta = any(scenario_delta == "ssp245"),
            `2000_rel` = any(period %in% c("2000s","1850s")),
            `2090_rel` = any(period == "2090s"))

sum(rel_doi$rel)/nrow(rel_doi)
sum(rel_doi$rel_delta)/nrow(rel_doi)
sum(rel_doi$rel_delta_co2)/nrow(rel_doi)

sum(rel_doi$`2000_rel`)/nrow(rel_doi)
sum(rel_doi$`2090_rel`)/nrow(rel_doi)

t <- subset(rel_doi,rel_delta == TRUE)
sum(t$ssp585_delta & !t$ssp245_delta)/nrow(t)
sum(t$ssp245_delta & !t$ssp585_delta)/nrow(t)
sum(t$ssp245_delta & t$ssp585_delta)/nrow(t)

(sum(t$ssp585_delta & !t$ssp245_delta)/nrow(t) +
sum(t$ssp245_delta & !t$ssp585_delta)/nrow(t) +
sum(t$ssp245_delta & t$ssp585_delta)/nrow(t))

OA_ICC$biome %>% table()


oa_icc_hits <- bind_rows(oa_icc_hits)
oa_icc_hits$p_s <-  paste(oa_icc_hits$scenario, oa_icc_hits$period,  sep = " - ")
oa_icc_hits[which(oa_icc_hits$p_s %in% c("ssp245 - 1850s","ssp585 - 1850s")),"p_s"] <- "Historical , 1850s"
oa_icc_hits[which(oa_icc_hits$p_s %in% c("ssp245 - 1850s","ssp585 - 1850s")),"p_s"] <- "Historical , 1850s"
oa_icc_hits[which(oa_icc_hits$p_s %in% c("ssp245 - 2000s","ssp585 - 2000s")),"p_s"] <- "Both, 2000s"
oa_icc_hits[which(oa_icc_hits$p_s %in% c("ssp245 - 2000s","ssp585 - 2000s")),"p_s"] <- "Both, 2000s"
oa_icc_hits[which(oa_icc_hits$p_s %in% c("ssp585 - 2090s")),"p_s"] <- "SSP5-8.5 , 2090s"
oa_icc_hits[which(oa_icc_hits$p_s %in% c("ssp245 - 2090s")),"p_s"] <- "SSP2-4.5 , 2090s"

nrow(oa_icc_hits)
table(oa_icc_hits %>% subset(p_s == "Historical , 1850s") %>% select(biome))
table(oa_icc_hits %>% subset(p_s == "Both, 2000s") %>% select(biome))
table(oa_icc_hits %>% subset(p_s == "SSP5-8.5 , 2090s") %>% select(biome))

biome_names <- c(
  "NP_ICE", "NP_SPSS", "NP_STSS", "NP_STPS",
  "Pac_EQU_W", "Pac_EQU_E", "SP_STPS", "NA_ICE",
  "NA_SPSS", "NA_STSS", "NA_STPS", "Atl_EQU",
  "SA_STPS", "IND_STPS", "SO_STSS", "SO_SPSS",
  "SO_ICE"
)

hits_2090s_ssp585 <- table(oa_icc_hits %>% subset(period == "2090s") %>% select(biome))
hits_2090s_ssp245 <- table(oa_icc_hits %>% subset(period == "2090s") %>% select(biome))
hits_2000s <- table(oa_icc_hits %>% subset(period == "2000s") %>% select(biome))
hits_2000s[biome_names[!biome_names %in% names(hits_2000s)]] <- 0

hits <- table(oa_icc_hits %>% select(biome))
hits[biome_names[!biome_names %in% names(hits)]] <- 0


base_loc <- "C:/Users/henryholm/OneDrive - Woods Hole Oceanographic Institution/Desktop/Galen_CMIP_Collab/raw_data/"

# Biome data
biomes <- terra::rast(paste0(base_loc, "Time_Varying_Biomes.nc"))

# Format Galen biome data
biomes <- flip(t(biomes))
crs(biomes) <- "+proj=longlat +ellips=WGS84"
biomes_vec <- as.polygons(biomes)
biomes_vec$MeanBiomes <- c(
  "NP_ICE", "NP_SPSS", "NP_STSS", "NP_STPS",
  "Pac_EQU_W", "Pac_EQU_E", "SP_STPS", "NA_ICE",
  "NA_SPSS", "NA_STSS", "NA_STPS", "Atl_EQU",
  "SA_STPS", "IND_STPS", "SO_STSS", "SO_SPSS",
  "SO_ICE"
)
biomes_vec$oa_icc_hits <- hits[c(
  "NP_ICE", "NP_SPSS", "NP_STSS", "NP_STPS",
  "Pac_EQU_W", "Pac_EQU_E", "SP_STPS", "NA_ICE",
  "NA_SPSS", "NA_STSS", "NA_STPS", "Atl_EQU",
  "SA_STPS", "IND_STPS", "SO_STSS", "SO_SPSS",
  "SO_ICE"
)]

options(terra.pal=grDevices::grey.colors(10,rev = T))
plot(biomes_vec, "oa_icc_hits")

biomes_rast <- rast(biomes_vec)

ggplot() +
  tidyterra::geom_spatvector(aes(fill = oa_icc_hits),data = biomes_vec) +
  theme_minimal()


# test our lab conditions
lab_cond <- lab %>% filter(samp_group %in% c("all_sample", "all_blank") & Name != "7002")

test <- list()
for (i in 1:nrow(lab_cond)) {
  test[[i]] <- (
    any(
    within_k_sd(as.numeric(lab_cond[i,"seacarb_CO2_mean"]*1e6), delta$areaW_mean_abs_aqueous, delta$sd_abs_aqueous) &
    within_k_sd(as.numeric(lab_cond[i, "seacarb_HCO3_mean"]*1e6), delta$areaW_mean_abs_bicarbonate, delta$sd_abs_bicarbonate) &
    within_k_sd(as.numeric(lab_cond[i, "seacarb_CO3_mean"]*1e6), delta$areaW_mean_abs_carbonate, delta$sd_abs_carbonate)  &
    within_k_sd(as.numeric(lab_cond[i,"Temp.nominal_mean"]), delta$areaW_mean_abs_tos, delta$sd_abs_tos) 
    )
  )
}

lab_cond$test <- unlist(test)

testing_studys <- OA_ICC %>%
  group_by(Study_DOI) %>%
  summarise(study_ocean_relevant = any(ocean_relevant))

sum(testing_studys$study_ocean_relevant)/nrow(testing_studys)

OA_ICC %>%
  ggplot() +
  geom_density(aes(CO2_mean),bw = 0.06) +
  theme_minimal() + lims(x = c(0,250)) + labs(x = co2,y = "Density") +
OA_ICC %>%
  subset(CO2_mean < 100 & HCO3_mean < 2600) %>%
  ggplot() +
  geom_density_2d_filled(aes(y = HCO3_mean, x = CO2_mean),bins = 90)+
  theme_minimal() + theme(legend.position = "none") +
  scale_fill_viridis_d(option = "H") +
  labs(x = co2, y = hco3) +
  OA_ICC %>%
  subset(CO2_mean < 100 & CO3_mean < 600) %>%
  ggplot() +
  geom_density_2d_filled(aes(y = CO3_mean, x = CO2_mean),bins = 90)+
  theme_minimal() + theme(legend.position = "none") +
  scale_fill_viridis_d(option = "H") +
  labs(x = co2, y = co3) + plot_layout(design = "AA\nBC")

ggsave("C:/Users/henryholm/OneDrive - Woods Hole Oceanographic Institution/Desktop/Galen_CMIP_Collab/Figures/Fig2_denisty_supp.pdf",
       device = cairo_pdf, width = 7, height = 7)

```

```{r}
delta %>%
  subset(type != "ICE" & p_s != "ssp585 - 2000s") %>%
  group_by(scenario, biome, loc, type, period) %>%
  ungroup() %>%
  ggplot(aes(y = areaW_mean_abs_tos, x = areaW_mean_abs_aqueous, color = p_s)) +
  #ggdensity::geom_hdr(aes(y = HCO3_mean, x = CO2_mean, color = NA),
  #                fill = 'grey70',data = OA_ICC) +
  geom_point(aes(y = Temp_mean, x = CO2_mean, color = NA),
                  color = 'grey70',alpha = 0.3,shape = 19, data = OA_ICC) +
  geom_point(aes(shape = type)) +
  ggdensity::geom_hdr_points(aes(y = Temp_mean, x = CO2_mean, color = NA),
                  color = 'grey70',alpha = 0.3,shape = 19, data = OA_ICC) +
  geom_errorbar(aes(xmin = areaW_mean_abs_aqueous - sd_abs_aqueous*2,
                    xmax = areaW_mean_abs_aqueous + sd_abs_aqueous*2)) +
  geom_errorbar(aes(ymin = areaW_mean_abs_tos - sd_abs_tos*2,
                    ymax = areaW_mean_abs_tos + sd_abs_tos*2)) +
 # labs(x = co2, y = hco3,color = "Period and Scenario",shape = "Biome Type") +
  scale_color_manual(values = c("grey20","orange","#3295a8","maroon")) +
  plot_layout(ncol = 1,guides = "collect") +
  theme_minimal() +
  scale_shape_manual(values = list("EQU" = 15, "STPS" = 16, "STSS" = 17, "SPSS" = 4)) +
  lims(x = c(0,250)) 
delta %>%
  subset(type != "ICE" & p_s != "ssp585 - 2000s") %>%
  group_by(scenario, biome, loc, type, period) %>%
  ungroup() %>%
  ggplot(aes(y = areaW_mean_abs_tos, x = areaW_mean_abs_aqueous, color = p_s)) +
  #ggdensity::geom_hdr(aes(y = HCO3_mean, x = CO2_mean, color = NA),
  #                fill = 'grey70',data = OA_ICC) +
  geom_point(aes(y = Temp_mean, x = CO2_mean, color = miss_cooccurrence),
                  shape = 19, data = OA_ICC) +
  geom_point(aes(shape = type)) +
  geom_errorbar(aes(xmin = areaW_mean_abs_aqueous - sd_abs_aqueous*2,
                    xmax = areaW_mean_abs_aqueous + sd_abs_aqueous*2)) +
  geom_errorbar(aes(ymin = areaW_mean_abs_tos - sd_abs_tos*2,
                    ymax = areaW_mean_abs_tos + sd_abs_tos*2)) +
 # labs(x = co2, y = hco3,color = "Period and Scenario",shape = "Biome Type") +
  scale_color_manual(values = c("black","grey20","orange","#3295a8","maroon","green")) +
  plot_layout(ncol = 1,guides = "collect") +
  theme_minimal() +
  scale_shape_manual(values = list("EQU" = 15, "STPS" = 16, "STSS" = 17, "SPSS" = 4)) +
  lims(x = c(0,100)) 

delta %>%
  subset(type != "ICE" & p_s != "ssp585 - 2000s") %>%
  group_by(scenario, biome, loc, type, period) %>%
  ungroup() %>%
  ggplot(aes(y = areaW_mean_abs_pH, x = areaW_mean_abs_aqueous, color = p_s)) +
  #ggdensity::geom_hdr(aes(y = HCO3_mean, x = CO2_mean, color = NA),
  #                fill = 'grey70',data = OA_ICC) +
  geom_point(aes(y = Temp_mean, x = CO2_mean, color = miss_cooccurrence),
                  shape = 19, data = OA_ICC) +
  geom_point(aes(shape = type)) +
  geom_errorbar(aes(xmin = areaW_mean_abs_aqueous - sd_abs_aqueous*2,
                    xmax = areaW_mean_abs_aqueous + sd_abs_aqueous*2)) +
  geom_errorbar(aes(ymin = areaW_mean_abs_tos - sd_abs_tos*2,
                    ymax = areaW_mean_abs_tos + sd_abs_tos*2)) +
 # labs(x = co2, y = hco3,color = "Period and Scenario",shape = "Biome Type") +
  scale_color_manual(values = c("black","grey20","orange","#3295a8","maroon","green")) +
  plot_layout(ncol = 1,guides = "collect") +
  theme_minimal() +
  scale_shape_manual(values = list("EQU" = 15, "STPS" = 16, "STSS" = 17, "SPSS" = 4)) +
  lims(x = c(0,100)) 

```

```{r}

dic %>% ggplot() + geom_histogram(aes(`DIC (mM)`))

joined %>%
  select(`DIC (mM)`, Incubator.CO2.ppm.nominal) %>%
  unique.data.frame() %>%
  ggplot() +
  geom_histogram(aes(`DIC (mM)`, fill = factor(Incubator.CO2.ppm.nominal))) +
  labs(fill = "Headspace pCO2", y = "Sample Count") +
  theme_minimal()

joined %>%
  select(`DIC (mM)`, Incubator.CO2.ppm.nominal) %>%
  unique.data.frame() %>%
  ggplot() +
  geom_density(aes(`DIC (mM)`, color = factor(Incubator.CO2.ppm.nominal)), size = 1) +
  labs(color = "Headspace pCO2", y = "Density") +
  theme_minimal()

```

