---
title: "R Notebook"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

load data and packages
```{r}

library(readxl) # read files 
library(reticulate) # use python packages 
library(tidyverse) # data formating / calculations 
library(here) # location management 
library(seacarb) # carbonate system calculations 
library(errors) # error calculation
library(patchwork) # plotting 
library(magrittr) # pipping 

# lab dic and ph measurements of final cultures 
ph <- read.csv(here("data/raw/ph.csv"))
dic <- read.csv(here("data/raw/dic.csv"))

# contains data on culture pH at variable cell densities to assess limits 
pHcd_all <- read.csv(here("data/raw/all_pH_training.csv"))

# annual biome data
biomes <- read.csv(here("data/product_data_processed/wide_annual_biomes_values.csv")) %>% as_tibble() 

# BATS data
BATS <- read.delim(here("data/raw/bats_bottle_formatted.txt"), header = T)

# set plot theme for everything 
ggplot2::theme_set(ggplot2::theme_minimal())

```

PyCO2Sys Calculations - Set up Python Env
```{r}

# create env
virtualenv_create("co2sys-env",python = "C:/Users/henryholm/AppData/Local/Python/pythoncore-3.14-64/python.exe")
use_virtualenv("co2sys-env", required = TRUE)
virtualenv_install("co2sys-env", "PyCO2SYS")

# import 
pyco2 <- import("PyCO2SYS")

```

Lab Data calculations - DIC and pH Calculate 
```{r}
# calculate pH using SOP 6a
for (i in seq_along(ph$pH_volt)) {
  ph[i, "pH_total_volt"] <- seacarb::pH(
    Ex = ph[i, "pH_volt"], # sample voltage
    Etris = ph[i, "pH_std_volt"], # std voltage
    T = ph[i, "pH_temp"], # temp
    S = 35 # salinity
  )
}

# join ph and dic data
lab_data <- left_join(dic,ph,join_by(Sample.ID.Offical == Culture_ID.),relationship = "many-to-many")

# add nominal temp data
lab_data <- lab_data %>% mutate(
  Temp.nominal = if_else(Incubator.CO2.ppm.nominal == 420, true = 22, false = NA),
  Temp.nominal = if_else(Incubator.CO2.ppm.nominal == 1135,
    true = 25.3, false = Temp.nominal
  )
)

### full carb calc
lab_data <- pyco2$sys(
  par1      = lab_data$pH_total_volt, # total ph
  par2      = lab_data$DIC_mM/1.0265/1e3, # DIC in μmol kg-1, measured density of media to get kg/L 
  par1_type = 3L, 
  par2_type = 2L,
  salinity  = 35,
  opt_pH_scale = 1, # total scale
  temperature = lab_data$Temp.nominal, # incubator temp 
  pressure  = 0,
  opt_k_carbonic = 10,
  opt_k_bisulfate = 1,
  opt_total_borate =1,
  opt_k_fluoride = 1
) %>% 
  data.frame() %>% 
  rename_with(~ paste0("pyco2sys_", .)) %>%
  cbind(lab_data,.)
  
lab_data <- seacarb::carbfull(flag = 9,
                        var1 = lab_data$pH_total_volt,
                        var2 = lab_data$DIC_mM/1.0265/1e3,
                        T = lab_data$Temp.nominal,
                        S = 35,
                        pHscale = "T",
                        k1k2 = "l", ks = "d", kf = "dg" # match PyCO2sys settings
                        ) %>% 
  rename_with(~ paste0("seacarb_", .)) %>%
  cbind(lab_data,.)

# identical - small floating point error diffs 
plot(lab_data$seacarb_CO2,lab_data$pyco2sys_CO2)
abline(b = 1,a = 0)
plot(lab_data$seacarb_HCO3,lab_data$pyco2sys_HCO3)
abline(b = 1,a = 0)

# final mean values 
lab_plot <- lab_data %>%
  group_by(Sample.ID.Offical) |> # group by sample
  subset(Blank %in% c("Sample_Final")) %>% # only look at final dic sample
  slice_max(Growth_Time_Nominal_hours, n = 1) %>% # only look at last pH meas
  group_by(Incubator.CO2.ppm.nominal) %>%
    summarise(
    across(
      where(is.numeric),
      .fns = list(
        mean = ~mean(unique(.x), na.rm = TRUE),
        sd   = ~sd(unique(.x), na.rm = TRUE)
      )
    )
  )%>%
  mutate(Name = "All Sp")

```

BATS Data Formatting
This chunk of code cuts 30 years of BATS data into a slice at 10 and 100 meters and takes averages.
```{r}

# id missing values as NA 
BATS[BATS == -999.000] <- NA

# refactor
BATS$DIC <- BATS$CO2
BATS$CO2 <- NULL

# calculate water pressure
BATS$hydrostatic_pressure_bar <- seacarb::d2p(BATS$Depth,lat = 32.166667)/10

# plot raw data
ggplot(BATS) + geom_point(aes(Temp,Depth)) + scale_y_reverse() 
ggplot(BATS) + geom_point(aes(Alk,Depth)) + scale_y_reverse() +
ggplot(BATS) + geom_point(aes(DIC,Depth)) + scale_y_reverse()

BATS$month <- stringr::str_remove(BATS$yyyymmdd,pattern = "^.{4}") %>%
  stringr::str_extract(pattern = "(.{2})")

# calculate co2 enviroment
sel <- BATS %>%
  select(Depth,Alk,DIC,Temp,Sal1,hydrostatic_pressure_bar) %>%
  complete.cases()

BATS_carb <- BATS %>%
  select(Depth,decy,month,Alk,DIC,Temp,Sal1,hydrostatic_pressure_bar,
         Pro,POC,Syn,Piceu,Naneu,yyyymmdd) %>%
  subset(sel)

BATS_carb1 <- BATS_carb %>% select(-Alk,-DIC) %>% cbind(.,BATS_carb %$%
          seacarb::carbb(flag = 15, var1 = Alk/1e6, var2 = DIC/1e6,
                        T = Temp, S = Sal1, 
                        P = hydrostatic_pressure_bar))


### full carb calc
BATS_carb <- pyco2$sys(
  par1      = BATS_carb$Alk,
  par2      = BATS_carb$DIC,
  par1_type = 1L,
  par2_type = 2L,
  salinity  = BATS_carb$Sal1,
  opt_pH_scale = 1,
  temperature = BATS_carb$Temp,
  pressure  = BATS_carb$hydrostatic_pressure_bar/10, #dbar
  opt_k_carbonic = 10,
  opt_k_bisulfate = 1,
  opt_total_borate =1,
  opt_k_fluoride = 1
) %>% 
  data.frame() %>% 
  rename_with(~ paste0("pyco2sys_", .)) %>%
  cbind(BATS_carb,.)

ggplot(BATS_carb) +
  geom_point(aes(pyco2sys_pCO2,Depth)) +
  scale_y_continuous(trans = scales::transform_compose(
    scales::transform_sqrt(), 
    scales::transform_reverse()
  ))

library(ggpointdensity)
BATS_carb %>%
  mutate(CO2 = pyco2sys_CO2*1e6,
         HCO3 = pyco2sys_HCO3*1e6,
         CO3 = pyco2sys_CO3*1e6,
         DIC = pyco2sys_dic*1e6,
         ALK = pyco2sys_alkalinity*1e6,
         pCO2 = pyco2sys_pCO2,
         pH = pyco2sys_pH_total) %>%
  pivot_longer(c("pCO2","pH","CO2","HCO3","CO3","DIC","ALK","Pro","POC","Syn","Piceu","Naneu","Temp"),names_to = "carb_var") %>%
  ggplot() + geom_point(aes(value,Depth)) +
  scale_y_reverse() + lims(y = c(1000,0)) +
  geom_hline(yintercept = 200,color = 'red') +
  facet_wrap(vars(carb_var),scales = "free_x")

# depth cut
BATS_dsub <- rbind(
  BATS_carb %>% subset(Depth > 95 & Depth < 105) %>% mutate(nominal_depth = 100),
  BATS_carb %>% subset(Depth > 35 & Depth < 45) %>% mutate(nominal_depth = 40),
  BATS_carb %>% subset(Depth > 7.5 & Depth < 12.5) %>% mutate(nominal_depth = 10),
  BATS_carb %>% subset(Depth < 5) %>% mutate(nominal_depth = 0)
  )

BATS_dsub %>%
  mutate(CO2 = pyco2sys_CO2*1e6, HCO3 = pyco2sys_HCO3*1e6, CO3 = pyco2sys_CO3*1e6,
         DIC = pyco2sys_dic*1e6,ALK = pyco2sys_alkalinity*1e6,pCO2 = pyco2sys_pCO2,
         pH = pyco2sys_pH_total) %>%
  pivot_longer(c("pCO2","pH","CO2","HCO3","CO3","DIC","ALK","Temp"),
               names_to = "carb_var") %>%
  ggplot() + geom_line(aes(decy,value,color = as.character(nominal_depth))) +
  theme(legend.position = "top",strip.placement = "outside") +
  labs(y = "",x = "Time (Years)",color = "Nominal Depth (m)") +
  facet_wrap(vars(carb_var),scales = "free_y",strip.position = "left",axes = 'all_x')

BATS_dsub <- BATS_dsub %>%
  mutate(CO2 = pyco2sys_CO2, HCO3 = pyco2sys_HCO3, CO3 = pyco2sys_CO3,
         DIC = pyco2sys_dic,ALK = pyco2sys_alkalinity,pCO2 = pyco2sys_pCO2,
         pH = pyco2sys_pH_total) 

t <- spline(y = BATS_dsub$CO2, x = BATS_dsub$decy)

ggplot() + geom_point(aes(x = BATS_dsub$decy,y = BATS_dsub$CO2)) + geom_line(aes(t$x,t$y))

# get month averages
BATS_dsub <- BATS_dsub %>%
  subset(nominal_depth %in% c(0,10,40,100)) %>%
  mutate(yyyymmdd = as.character(yyyymmdd)) %>%
  separate(yyyymmdd,into = c("year","mmdd"),sep = 4) %>%
  separate(mmdd,into = c("month","day"),sep = 2) %>%
  group_by(year,month,nominal_depth) %>%
  summarise_if(.predicate = is.numeric,.funs = ~ mean(.x, na.rm = T))

BATS_dsub$decade <- cut(BATS_dsub$decy,breaks = c(1990,2000,2010,2020,2030)) %>%
  fct_recode(`1990s` = "(1.99e+03,2e+03]",
             `2000s` = "(2e+03,2.01e+03]",
             `2010s` = "(2.01e+03,2.02e+03]",
             `2020s` = "(2.02e+03,2.03e+03]")

BATS_dsub$year <- cut(BATS_dsub$decy,breaks = round(min(BATS_dsub$decy)):ceiling(max(BATS_dsub$decy))) 

BATS_dsub_summary <- BATS_dsub %>%
  group_by(decade,nominal_depth) %>%
    summarise(across(
    .cols = c(Temp, DIC,pH,CO2,pCO2,HCO3,CO3,ALK),  # Add your columns here
    .fns = list(mean = ~mean(.x, na.rm = TRUE),
                sd = ~sd(.x, na.rm = TRUE)),
    .names = "{.col}_{.fn}"
  ))

lm(Temp ~ decy, BATS_dsub %>% subset(nominal_depth == 0 & decy < 2019))

BATS_dsub_summary_y <- BATS_dsub %>%
  group_by(year,nominal_depth) %>%
    summarise(across(
    .cols = c(Temp, DIC,pH,CO2,pCO2,HCO3,CO3,ALK),  # Add your columns here
    .fns = list(mean = ~mean(.x, na.rm = TRUE),
                sd = ~sd(.x, na.rm = TRUE)),
    .names = "{.col}_{.fn}"
  ))

BATS_dsub %>% subset(nominal_depth %in% c(0,40,100)) %>% ggplot() + geom_point(aes(month,Temp,color = factor(nominal_depth)))

BATS_dsub %>% subset(nominal_depth %in% c(0,40,100)) %>% ggplot() + geom_point(aes(month,HCO3,color = factor(nominal_depth)))

# clean up
BATS <- list(dataframe = BATS_dsub, means = BATS_dsub_summary, year_means = BATS_dsub_summary_y)
remove(BATS_carb,BATS_dsub,BATS_dsub_summary,sel,t,a)
gc()

```

pH cell density limit panel
```{r}

# convert pvalue to significance level
signif.num <- function(x) {
    symnum(x, corr = FALSE, na = FALSE, legend = FALSE,
           cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1), 
           symbols = c("<0.001", "<0.01", "<0.05", "<0.1", "NonSig"))
}

sp <- c("Ehux","Tpsu")
formal <- c(Ehux = "G. huxleyi", Tpsu = "T. pseudonana")

ph_cd <- list()

# dfs
for (i in 1:length(sp)) {
  ph_cd$dfs[[sp[i]]] <- pHcd_all %>%
    subset(Species == sp[i] &
      complete.cases(.))
}

# models
for (i in 1:length(sp)) {
ph_cd$exp_mods[[sp[i]]]<- ph_cd$dfs[[sp[i]]] %>%
  mutate(lnCellDensity = log10(CellDensity), lnpH_offset = log10(pH_offset)) %>%
  lm(lnpH_offset ~ lnCellDensity,data = .)
}

ph_cd_param <- map_dfr(sp, function(species) {
  mod <- ph_cd[["exp_mods"]][[species]]
  mod_summary <- summary(mod)
  
  tibble(
    Species = species,
    Intercept = coef(mod)[1],
    Slope = coef(mod)[2],
    `Adj. R^2` = mod_summary$adj.r.squared,
    `Slope p-value` = signif.num(mod_summary$coefficients[2, 4]),
    `Intercept p-value` = signif.num(mod_summary$coefficients[1, 4]),
  )
})

ph_cd_param

ph <- seq(0,2,0.01)
names(ph_cd_param$Intercept) <- ph_cd_param$Species
ph <- cbind(ph,as.data.frame(map2(ph_cd_param$Intercept,ph_cd_param$Slope,~ 10^((log10(ph) - .x)/.y)))) %>%
  pivot_longer(ph_cd_param$Species,names_to = 'sp',values_to = "cd")
  
for (i in 1:length(sp)) {
  ph_offset <- 0.1
  cells_ml <- as.numeric(ph %>% subset(sp == sp[i] & ph == ph_offset) %>% select(cd))
  ph_cd[["plots"]][[sp[i]]] <- ph_cd$dfs[[sp[i]]] %>%
  ggplot() +
  geom_point(aes(CellDensity,pH_offset,
                 #color = container,shape = as.character(Incubator.CO2.ppm.nominal)
                 ),size = 2) +
  geom_line(aes(cd,ph),data = subset(ph,sp == sp[i]),size = 0.5) +
  annotate(geom = "segment",y = ph_offset,yend = ph_offset,x = 0,xend = cells_ml,linetype = "dashed") +
  annotate(geom = "segment",y = -0.2,yend = ph_offset,x = cells_ml,xend = cells_ml,linetype = "dashed") +
  scale_x_log10() +
  scale_y_continuous(limits = c(-0.2,1.75),expand = expansion()) +
  annotate("text", y = 1.4, x = min(ph_cd$dfs[[sp[i]]]$CellDensity, na.rm = TRUE), size = 4, fontface = 'italic', 
         label = formal[sp[i]], hjust = 0, vjust = 1.1) +
  ggsci::scale_color_npg() +
  labs(x = "log10 Cell Density (cells mL-1)", y = "ΔpH",color = "Container",shape = "Air CO2 Level") +
  ggpubr::theme_pubclean() + theme(legend.position = "right")

}

cells_ml_ehux <- as.numeric(ph %>% subset(sp == "Ehux" & ph == ph_offset) %>% select(cd))
cells_ml_tpsu <- as.numeric(ph %>% subset(sp == "Tpsu" & ph == ph_offset) %>% select(cd))

pH_panel <- ph_cd$dfs %>%
  bind_rows() %>%
  ggplot() +
  geom_point(aes(CellDensity,pH_offset,
                 color = Species,
                 ),size = 2) +
  geom_line(aes(cd,ph,color = "Ehux"),data = subset(ph,sp == "Ehux"),size = 1) +
  geom_line(aes(cd,ph,color = "Tpsu"),data = subset(ph,sp == "Tpsu"),size = 1) +
  annotate(geom = "segment",y = ph_offset,yend = ph_offset,x = 0,xend = cells_ml_ehux,linetype = "dashed") +
  annotate(geom = "segment",y = -0.2,yend = ph_offset,x = cells_ml_ehux,xend = cells_ml_ehux,linetype = "dashed") +
  annotate(geom = "segment",y = ph_offset,yend = ph_offset,x = 0,xend = cells_ml_tpsu,linetype = "dashed") +
  annotate(geom = "segment",y = -0.2,yend = ph_offset,x = cells_ml_tpsu,xend = cells_ml_tpsu,linetype = "dashed") +
  scale_x_log10(breaks = 10^(3:6)) +
  scale_y_continuous(limits = c(-0.2,1.75),expand = expansion()) +
  scale_color_manual(values = c('darkgreen','#5cd65c')) +
  labs(x = "Cell Density (log10 cells/mL)", y = "ΔpH",shape = "Air CO2 Level") +
  ggpubr::theme_pubclean() + theme(legend.position = "top")

pH_panel

```

Supp Figure Ghux co2 and container level
```{r}
ph_cd$dfs$Ehux %>%
  mutate(container = str_replace(container,pattern = "150rpm",replacement = "150 rpm"),
         container = str_replace(container,pattern = "125ml", replacement = "125ml flasks")) %>%
  
  ggplot() +
  geom_point(aes(CellDensity,pH_offset,
                 color = container, shape = as.factor(Incubator.CO2.ppm.nominal),
                 ),size = 3) +
  geom_line(aes(cd,ph),data = subset(ph,sp == "Ehux"),size = 1) +
  annotate(geom = "segment",y = ph_offset,yend = ph_offset,x = 0,xend = cells_ml_ehux,linetype = "dashed") +
  annotate(geom = "segment",y = -0.2,yend = ph_offset,x = cells_ml_ehux,xend = cells_ml_ehux,linetype = "dashed") +
  scale_x_log10(breaks = 10^(3:6)) +
  #scale_y_continuous(limits = c(-0.2,1.75),expand = expansion()) +
  labs(x = "Cell Density (log10 cells/mL)", y = "ΔpH",
       shape = expression(Air ~ CO[2] ~ Level), color = "Container and Shaking Speed") +
  ggpubr::theme_pubclean() + theme(legend.position = "left")
```


Figure 4 - DIC Alignment with NA and lab replication
```{r}

#### Condition Panel, north atlantic
choose_biomes <- c("NA_STPS","NA_STSS","Atl_EQU")

probs <- c(0.95, 0.80, 0.5)

vars <- colnames(lab_plot)[str_detect(colnames(lab_plot),"pyco2sys_")] %>%
  str_remove_all("_mean") %>% str_remove_all("_sd") %>% unique()

# refactor function
refactor_vars <- function(df, dataset, x_var, y_var, ymult = 1,xmult = 1, keep = NULL) {
  x_mean <- sym(paste0(x_var, "_mean"))
  x_sd   <- sym(paste0(x_var, "_sd"))
  y_mean <- sym(paste0(y_var, "_mean"))
  y_sd   <- sym(paste0(y_var, "_sd"))
  
  df %>%
    transmute(
      dataset = dataset,
      x    = !!x_mean * xmult,
      y    = !!y_mean * ymult,
      x_sd = !!x_sd   * xmult,
      y_sd = !!y_sd   * ymult,
      !!!syms(keep)
    )
}

# func to grap multiple datasets and group w bind_rows
refactor_many <- function(specs) {
  map_dfr(specs, function(s) {
    refactor_vars(
      df      = s$df,
      dataset = s$dataset,
      x_var   = s$x_var,
      y_var   = s$y_var,
      xmult    = s$xmult %||% 1,
      ymult    = s$ymult %||% 1,
      keep    = s$keep %||% NULL
    )
  })
}


# CMIP6 pre-processing step
cmip6_wide <- biomes %>%
  subset(biome %in% choose_biomes) 

levels_order <- c(
  "Historical , 1850s",
  "Modern , 2000s",
  "SSP2-4.5 , 2090s",
  "SSP5-8.5 , 2090s",
  "BATS",
  "Lab Data, this study"
)

cmip6_wide$p_s <- factor(cmip6_wide$p_s,levels = levels_order)

cmip6_wide <- cmip6_wide %>% rename(aqueous_mean = "areaW_mean_delta_aqueous",
                      aqueous_sd = "sd_delta_aqueous",
                      bicarbonate_mean = "areaW_mean_delta_bicarbonate",
                      bicarbonate_sd = "sd_delta_bicarbonate",
                      carbonate_mean = "areaW_mean_delta_carbonate",
                      carbonate_sd = "sd_delta_carbonate",
                      tos_mean = "areaW_mean_delta_tos",
                      tos_sd = "sd_delta_tos")

# Define the list of datasets + specs
dat1 <- list(
  list(df = subset(BATS$means, nominal_depth == 0 & decade != "2020s"),
       dataset = "BATS", y_var = "CO2", x_var = "HCO3", ymult = 1, xmult = 1),
  list(df = cmip6_wide, dataset = "CMIP6", y_var = "aqueous", x_var = "bicarbonate", keep = c("p_s","type")),
  list(df = lab_plot, dataset = "Lab Data, this study", y_var = "pyco2sys_CO2", x_var = "pyco2sys_HCO3",
       ymult = 1e6,xmult = 1e6, keep = c("Incubator.CO2.ppm.nominal"))
) %>%
  refactor_many() %>%
  mutate(#samp_group = if_else(is.na(samp_group), "", samp_group),
    p_s = if_else(is.na(p_s), dataset, p_s),
    p_s = factor(p_s,levels_order),
    shape = type,
    shape = if_else(is.na(shape),dataset,shape))

errors(dat1$x) <- dat1$x_sd
errors(dat1$y) <- dat1$y_sd

baseline_lab_x <- dat1[which(dat1$Incubator.CO2.ppm.nominal == 420),] %>% pull(x)
baseline_lab_y <- dat1[which(dat1$Incubator.CO2.ppm.nominal == 420),] %>% pull(y)

baseline_BATS_x <- dat1[which(dat1$decade == "2000s"),] %>% pull(x)
baseline_BATS_y <- dat1[which(dat1$decade == "2000s"),] %>% pull(y)


dat1 <- dat1 %>%
  ungroup() %>%
  mutate(
    x = if_else(
      dataset == "Lab Data, this study",
      x - baseline_lab_x,
      x
    ),
    y = if_else(
      dataset == "Lab Data, this study",
      y - baseline_lab_y,
      y
    ),
    x = if_else(
      dataset == "BATS",
      x - baseline_BATS_x,
      x
    ),
    y = if_else(
      dataset == "BATS",
      y - baseline_BATS_y,
      y
    )
  )

dat1$x_sd <- errors(dat1$x) 
dat1$y_sd <- errors(dat1$y)

A <- dat1 %>% subset(x != 0 & dataset != "BATS") %>%
  ggplot(aes(x, y)) +
  geom_point(aes(color = p_s, shape = shape), size = 2, na.rm = TRUE) +
  geom_errorbar(aes(ymin = y - y_sd, ymax = y + y_sd, color = p_s), width = 0) +
  geom_errorbar(aes(xmin = x - x_sd, xmax = x + x_sd, color = p_s), width = 0) +
  labs(x = expression("Δ [HCO"[3]*"] (umol kg"^{-1}*")"),
       y = expression("Δ [CO"[2]*"] (umol kg"^{-1}*")")) +
  scale_color_manual(values = c("grey50","orange","maroon","darkgreen")) +
  scale_shape_manual(values = c("STPS" = 19,"STSS" = 17, "EQU" = 15,"Lab Data, this study" = 22, "BATS" = 10)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_line(aes(x, y,color = "BATS"),data = subset(dat1,dataset == "BATS"),color = "blue") +
  geom_point(aes(x, y,color = "BATS"),data = subset(dat1,dataset == "BATS"),color = "blue") +
theme(legend.position = "none") 

A

##### temp and co2 (panel B)

# Define the list of datasets + specs
dat2 <- list(
  list(df = subset(BATS$means, nominal_depth == 0 & decade != "2020s"), dataset = "BATS", y_var = "CO2", x_var = "Temp", ymult = 1, xmult = 1),
  list(df = cmip6_wide, dataset = "CMIP6", y_var = "aqueous", x_var = "tos", keep = c("p_s","type")),
  list(df = lab_plot, dataset = "Lab Data, this study", y_var = "pyco2sys_CO2", x_var = "Temp.nominal",
       ymult = 1e6,xmult = 1, keep = c("Incubator.CO2.ppm.nominal"))
) %>%
  refactor_many() %>%
  mutate(#samp_group = if_else(is.na(samp_group), "", samp_group),
    p_s = if_else(is.na(p_s), dataset, p_s),
    p_s = factor(p_s,levels_order),
    shape = type,
    shape = if_else(is.na(shape),dataset,shape))

errors(dat2$x) <- dat2$x_sd
errors(dat2$y) <- dat2$y_sd

baseline_lab_x <- dat2[which(dat1$Incubator.CO2.ppm.nominal == 420),] %>% pull(x)
baseline_lab_y <- dat2[which(dat1$Incubator.CO2.ppm.nominal == 420),] %>% pull(y)

baseline_BATS_x <- dat2[which(dat1$decade == "2000s"),] %>% pull(x)
baseline_BATS_y <- dat2[which(dat1$decade == "2000s"),] %>% pull(y)


dat2 <- dat2 %>%
  mutate(
    x = if_else(
      dataset == "Lab Data, this study",
      x - baseline_lab_x,
      x
    ),
    y = if_else(
      dataset == "Lab Data, this study",
      y - baseline_lab_y,
      y
    ),
    x = if_else(
      dataset == "BATS",
      x - baseline_BATS_x,
      x
    ),
    y = if_else(
      dataset == "BATS",
      y - baseline_BATS_y,
      y
    )
  )

dat2$x_sd <- errors(dat2$x) 
dat2$y_sd <- errors(dat2$y)

B <- dat2 %>% subset(x != 0 & dataset != "BATS") %>%
  ggplot(aes(x, y)) +
  geom_point(aes(color = p_s, shape = shape), size = 2, na.rm = TRUE) +
  geom_errorbar(aes(ymin = y - y_sd, ymax = y + y_sd, color = p_s), width = 0) +
  geom_errorbar(aes(xmin = x - x_sd, xmax = x + x_sd, color = p_s), width = 0) +
  labs(x = expression("Δ Temperature (C)"),
       y = expression("Δ [CO"[2]*"] (umol kg"^{-1}*")")) +
  scale_color_manual(values = c("grey50","orange","maroon","darkgreen")) +
  scale_shape_manual(values = c("STPS" = 19,"STSS" = 17, "EQU" = 15,"Lab Data, this study" = 22, "BATS" = 10)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_line(aes(x, y),data = subset(dat2,dataset == "BATS"),color = "blue") +
  geom_point(aes(x, y),data = subset(dat2,dataset == "BATS"),color = "blue") +
  theme(legend.position = "none") 

A + theme_minimal() + B + theme_minimal() + plot_layout(guides = "collect")

pretty_sci <- function(x) {
  parse(text = gsub("e\\+?", " %*% 10^", format(x, scientific = TRUE)))
}

pH_panel + theme(axis.title = element_text(size = 10)) + theme(legend.position = "none") +
  plot_spacer() +
  A + theme(axis.title = element_text(size = 10)) +
  B + theme(axis.title = element_text(size = 10)) + theme(legend.position = "none") +
  plot_layout(ncol = 2, axes = "collect",guides = "collect") +
  theme(
    plot.tag = element_text(
      face = "bold",
      size = 14,
      hjust = -0.1,
      vjust = 1.1
    )
  )

```

Non-delta version for supplement
```{r}

############### abs version for suppliment

# CMIP6 pre-processing step
cmip6_wide <- biomes %>%
  subset(biome %in% choose_biomes) 

cmip6_wide <- cmip6_wide %>% rename(aqueous_mean = "areaW_mean_abs_aqueous",
                      aqueous_sd = "sd_abs_aqueous",
                      bicarbonate_mean = "areaW_mean_abs_bicarbonate",
                      bicarbonate_sd = "sd_abs_bicarbonate",
                      carbonate_mean = "areaW_mean_abs_carbonate",
                      carbonate_sd = "sd_abs_carbonate",
                      pH_mean = "areaW_mean_abs_pH",
                      pH_sd = "sd_abs_pH",
                      tos_mean = "areaW_mean_abs_tos",
                      tos_sd = "sd_abs_tos")
# Define the list of datasets + specs
dat1 <- list(
  list(df = subset(BATS$means, nominal_depth == 0 & decade != "2020s"), dataset = "BATS", y_var = "CO2", x_var = "HCO3", ymult = 1, xmult = 1),
  list(df = cmip6_wide, dataset = "CMIP6", y_var = "aqueous", x_var = "bicarbonate", keep = c("p_s","type")),
  list(df = lab_plot, dataset = "Lab Data, this study", y_var = "pyco2sys_CO2", x_var = "pyco2sys_HCO3",
       ymult = 1e6,xmult = 1e6, keep = c("Incubator.CO2.ppm.nominal"))
) %>%
  refactor_many() %>%
  mutate(#samp_group = if_else(is.na(samp_group), "", samp_group),
    p_s = if_else(is.na(p_s), dataset, p_s),
    p_s = factor(p_s,levels_order),
    shape = type,
    shape = if_else(is.na(shape),dataset,shape))


A <- dat1 %>% # subset(dataset != "BATS") %>%
  ggplot(aes(x, y)) +
  geom_point(aes(color = p_s, shape = shape), size = 2, na.rm = TRUE) +
  geom_errorbar(aes(ymin = y - y_sd*2, ymax = y + y_sd*2, color = p_s), width = 0) +
  geom_errorbar(aes(xmin = x - x_sd*2, xmax = x + x_sd*2, color = p_s), width = 0) +
  labs(x = expression("[HCO"[3]*"] (umol kg"^{-1}*")"),
       y = expression("[CO"[2]*"] (umol kg"^{-1}*")")) +
  scale_color_manual(values = c("grey50","#3295a8","orange","maroon","blue","darkgreen")) +
  scale_shape_manual(values = c("STPS" = 19,"STSS" = 17,"EQU" = 15, "Lab Data, this study" = 22,"BATS" = 2)) +
  theme(legend.position = "none") 

# Define the list of datasets + specs
dat2 <- list(
  list(df = subset(BATS$means, nominal_depth == 0 & decade != "2020s"), dataset = "BATS", y_var = "CO2", x_var = "Temp", ymult = 1, xmult = 1),
  list(df = cmip6_wide, dataset = "CMIP6", y_var = "aqueous", x_var = "tos", keep = c("p_s","type")),
  list(df = lab_plot, dataset = "Lab Data, this study", y_var = "pyco2sys_CO2", x_var = "Temp.nominal",
       ymult = 1e6,xmult = 1, keep = c("Incubator.CO2.ppm.nominal"))
) %>%
  refactor_many() %>%
  mutate(#samp_group = if_else(is.na(samp_group), "", samp_group),
    p_s = if_else(is.na(p_s), dataset, p_s),
    p_s = factor(p_s,levels_order),
    shape = type,
    shape = if_else(is.na(shape),dataset,shape))

B <- dat2 %>% #subset(x != 0 & dataset != "BATS") %>%
  ggplot(aes(x, y)) +
  geom_point(aes(color = p_s, shape = shape), size = 2, na.rm = TRUE) +
  geom_errorbar(aes(ymin = y - y_sd*2, ymax = y + y_sd*2, color = p_s), width = 0) +
  geom_errorbar(aes(xmin = x - x_sd*2, xmax = x + x_sd*2, color = p_s), width = 0) +
  labs(x = expression("Temperature (C)"),
       y = expression("[CO"[2]*"] (umol kg"^{-1}*")")) +
  scale_color_manual(values = c("grey50","#3295a8","orange","maroon","blue","darkgreen")) +
  scale_shape_manual(values = c("STPS" = 19,"STSS" = 17,"EQU" = 15, "Lab Data, this study" = 22,"BATS" = 2)) +
  geom_line(aes(x, y),data = subset(dat2,dataset == "BATS"),color = "blue") +
  geom_point(aes(x, y),data = subset(dat2,dataset == "BATS"),color = "blue") +
  theme(legend.position = "none") 

# Define the list of datasets + specs
dat3 <- list(
  list(df = subset(BATS$means, nominal_depth == 0 & decade != "2020s"), dataset = "BATS", y_var = "CO2", x_var = "CO3", ymult = 1, xmult = 1),
  list(df = cmip6_wide, dataset = "CMIP6", y_var = "aqueous", x_var = "carbonate", keep = c("p_s","type")),
  list(df = lab_plot, dataset = "Lab Data, this study", y_var = "pyco2sys_CO2", x_var = "pyco2sys_CO3",
       ymult = 1e6,xmult = 1e6, keep = c("Incubator.CO2.ppm.nominal"))
) %>%
  refactor_many() %>%
  mutate(#samp_group = if_else(is.na(samp_group), "", samp_group),
    p_s = if_else(is.na(p_s), dataset, p_s),
    p_s = factor(p_s,levels_order),
    shape = type,
    shape = if_else(is.na(shape),dataset,shape))

C <- dat3 %>% # subset(dataset != "BATS") %>%
  ggplot(aes(x, y)) +
  geom_point(aes(color = p_s, shape = shape), size = 2, na.rm = TRUE) +
  geom_errorbar(aes(ymin = y - y_sd*2, ymax = y + y_sd*2, color = p_s), width = 0) +
  geom_errorbar(aes(xmin = x - x_sd*2, xmax = x + x_sd*2, color = p_s), width = 0) +
  labs(x = expression("[CO"[3]*"] (umol kg"^{-1}*")"),
       y = expression("[CO"[2]*"] (umol kg"^{-1}*")")) +
  scale_color_manual(values = c("grey50","#3295a8","orange","maroon","blue","darkgreen")) +
  scale_shape_manual(values = c("STPS" = 19,"STSS" = 17,"EQU" = 15, "Lab Data, this study" = 22,"BATS" = 2)) +
  theme(legend.position = "none") 


# Define the list of datasets + specs
dat4 <- list(
  list(df = subset(BATS$means, nominal_depth == 0 & decade != "2020s"), dataset = "BATS", y_var = "CO2", x_var = "pH", ymult = 1, xmult = 1),
  list(df = cmip6_wide, dataset = "CMIP6", y_var = "aqueous", x_var = "pH", keep = c("p_s","type")),
  list(df = lab_plot, dataset = "Lab Data, this study", y_var = "pyco2sys_CO2", x_var = "pyco2sys_pH_total",
       ymult = 1e6,xmult = 1, keep = c("Incubator.CO2.ppm.nominal"))
) %>%
  refactor_many() %>%
  mutate(#samp_group = if_else(is.na(samp_group), "", samp_group),
    p_s = if_else(is.na(p_s), dataset, p_s),
    p_s = factor(p_s,levels_order),
    shape = type,
    shape = if_else(is.na(shape),dataset,shape))

D <- dat4 %>% # subset(dataset != "BATS") %>%
  ggplot(aes(x, y)) +
  geom_point(aes(color = p_s, shape = shape), size = 2, na.rm = TRUE) +
  geom_errorbar(aes(ymin = y - y_sd*2, ymax = y + y_sd*2, color = p_s), width = 0) +
  geom_errorbar(aes(xmin = x - x_sd*2, xmax = x + x_sd*2, color = p_s), width = 0) +
  labs(x = "pH (Total Scale)",
       y = expression("[CO"[2]*"] (umol kg"^{-1}*")")) +
  scale_color_manual(values = c("grey50","#3295a8","orange","maroon","blue","darkgreen")) +
  scale_shape_manual(values = c("STPS" = 19,"STSS" = 17,"EQU" = 15, "Lab Data, this study" = 22,"BATS" = 2)) +
  theme(legend.position = "none") 

A + B + C + D + theme(legend.position = "right") 


```

